---
title: "MUPOLYGON Report"
author: "`r Sys.getenv('USERNAME')`"
date: "`r Sys.Date()`"
output: 
  html_document: 
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
params:
   projectname: !r as.character("EVAL - MLRA 95 - Sisson fine sandy loam, 0 to 2 percent slopes;EVAL - MLRA 95 - Sisson fine sandy loam, 12 to 20 percent slopes, eroded;EVAL - MLRA 95 - Sisson fine sandy loam, 2 to 6 percent slopes;EVAL - MLRA 95 - Sisson fine sandy loam, 2 to 6 percent slopes, eroded;EVAL - MLRA 95 - Sisson fine sandy loam, 6 to 12 percent slopes, eroded")
   geodatabase:  !r as.character("RTSD_MLRA_11-JUE_FY18.gdb")
   project_data: !r as.character("M:/geodata/project_data/")
   mlrassoarea:  !r as.character("11-JUE")
---
# Report Description

This report summarizes the geographic setting of a list of musym within a shapefile. It is intented to be used to compare and contrast map units, and suggest possible Low, RV, and High values for soil components. 


```{r LoadPackages, include=FALSE}
options(stringsAsFactors = FALSE)
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# report parameters
source("config.R")

# load packages
suppressWarnings( {
  library(plyr)
  library(reshape2)
  library(DT)
  library(lattice)
  library(latticeExtra)# library(ggplot2)
  library(circular)
  library(maps)
  library(maptools)
  library(gdalUtils)
  library(sp)
  library(sf)
  library(raster)
  library(soilDB)
  library(soilReports)
  })

opar <- trellis.par.get()
trellis.par.set(theme = ggplot2like())
tpg <- trellis.par.get()
tpg$axis.line <-  list(alpha = 1, col = "#000000", lty = rep(1, 7), lwd = rep(0.2, 7))
tpg$strip.border <- list(alpha = 1, col = "#000000", lty = rep(1, 7), lwd = rep(0.2, 7))
trellis.par.set(tpg)

gdalUtils::gdal_setInstallation(search_path="C:/Program Files/QGIS Essen/bin", rescan = TRUE)
```


```{r map unit summary function}

# report parameters
attach(params)
projectname   <- unlist(strsplit(params$projectname, ";"))
mlrassoarea   <- gsub("-", "", mlrassoarea)
samplePercent <- 15
p             <- c(0, 0.25, 0.5, 0.75, 1)

# project data
project    <- get_projectmapunit_from_LIMS(projectname, stringsAsFactors = TRUE)
projectiid <- paste0(unique(project$projectiid), collapse = "_")
  
# folder paths
office_folder <- paste0(project_data, mlrassoarea, "/")
region_folder <- paste0(project_data, "11REGION/")

# cache files
sample_cache <- paste0(office_folder, "cache/", projectiid, "_", samplePercent, ".Rdata")
gdb_dsn   <- paste0(office_folder, geodatabase)
shp_dsn   <- paste0(office_folder, "cache/projectiid_", projectiid, ".shp")
gdb_cache <- gsub(".gdb", "_cache.RData", gdb_dsn)
shp_cache <- paste0(shp_dsn, ".RData")
```

```{r}
## local functions
if (!file.exists(sample_cache)){
  
  # extract shapefile via gdalUtils if the geodatabase is > 1.5GB
  if (
    # geodatabase is > 1.5GB
    sum(file.size(list.files(gdb_dsn, full.names = TRUE))) / 1e+9 > 1.5 & 
    # QGIS is installed
    file.exists("C:/Program Files/QGIS Essen/bin") &
    # shapefile of project doesn't already exist
    !file.exists(shp_cache) &
    # geodatabase cache doesn't exist
    !file.exists(gdb_cache)
    ){
    # extract polygons from geodatabase
    ogr_extract(office_folder, gdb_dsn, shp_dsn, project$lmapunitiid)
    mupolygon <- read_sf(dsn = shp_dsn, layer = paste0("projectiid_", projectiid))
    st_crs(mupolygon) <- "+init=epsg:5070"

    sapolygon <- read_sf(dsn = gdb_dsn, layer = "SAPOLYGON")
    st_crs(sapolygon) <- "+init=epsg:5070"

    save(mupolygon, sapolygon, file = shp_cache)
    }
  
  # shapefile cache
  if (file.exists(shp_cache)) load(file = shp_cache)
  
  # geodatabase cache
  if (
    # shapefile cache exist?
    !file.exists(shp_cache) &
    # geodatabase cache exists?
    !file.exists(gdb_cache) &
    # geodatabase size
    sum(file.size(list.files(gdb_dsn, full.names = TRUE))) / 1e+9 <= 1.5 
    ) {
    sapolygon <- read_sf(dsn = gdb_dsn, layer = "SAPOLYGON")
    st_crs(sapolygon) <- "+init=epsg:5070"
    
    mupolygon <- read_sf(dsn = gdb_dsn, layer = "MUPOLYGON")
    st_crs(mupolygon) <- "+init=epsg:5070"
    
    save(mupolygon, sapolygon, file = gdb_cache)
    }
  
  # geodatabase cache
  if (file.exists(gdb_cache)) load(file = gdb_cache)
  
  idx <- ! names(mupolygon) %in% c("Shape", "Geometry")
  names(mupolygon)[idx] <- tolower(names(mupolygon)[idx])
  idx2 <- grepl("shape_leng", names(mupolygon))
  names(mupolygon)[idx2] <- "shape_leng"
  
  mupolygon <- mupolygon[mupolygon$mukey %in% project$lmapunitiid, ]
  
  mupolygon <- within(mupolygon, {
    Acres = shape_area * 0.000247105
    Circularity = shape_leng / (2 * sqrt(shape_area / pi) * pi)
    })
  
  # Sample soil map unit
  # n <- ceiling(sum(mupolygon$shape_area) / 900 * samplePercent / 100)
  # 
  # mupolygon$idx <- 1:nrow(mupolygon)
  # mupolygon_sp <- by(mupolygon, mupolygon$idx, function(x) as(st_sample(x, 10), "Spatial"))
  # mupolygon_sp <- do.call(rbind, mupolygon_sp)
  # proj4string(mupolygon_sp) <- CRS("+init=epsg:5070")
  # mupolygon_sf <- st_as_sf(mupolygon_sp)
  
  mupolygons <- within(mupolygon, {
    idx    = 1:nrow(mupolygon)
    acres  = shape_area * 0.000247
    samp_n = ifelse(acres < 2, 1, round(acres / 2))
    })
  
  mupolygon_sp <- {
    split(mupolygons, mupolygons$idx) ->.;
    lapply(., function(x) {
      mu_sp = spsample(as(x, "Spatial"), x$samp_n, type = "random", iter = 10)
      if ("y" %in% names(as.data.frame(mu_sp))) {
        test = data.frame(as.data.frame(mu_sp), idx = x$idx[1])
        coordinates(test) = ~ x + y
        proj4string(test) = "+init=epsg:5070"
        return(test)
        }
      })}
  idx <- which(unlist(lapply(mupolygon_sp, function(x) !is.null(x))))
  mupolygon_sp <- mupolygon_sp[idx]
  mupolygon_sp <- do.call("rbind", mupolygon_sp)
  mupolygon_sf <- st_as_sf(mupolygon_sp)

  
  geodata <- raster_extract(mupolygon_sp)
  idx  <- st_intersects(mupolygon_sf, mupolygon)
  mapunit.df <- mupolygon[unlist(lapply(idx, function(x) x[1])), ]

  data <- cbind(data.frame(mapunit.df[c("areasymbol", "musym")]), geodata)
  
#  mlra <- readOGR(dsn = "M:/geodata/project_data/11REGION/mlra_a_r11.shp", layer = "mlra_a_r11", encoding = "ESRI Shapefile")
#  mlra_i <- over(mupolygon_sp, mlra)
#  mlra_acres <- ddply(mlra_i, .(MLRARSYM), summarize, mlra_percent = sum(length(OBJECTID)))
#  mlra_acres <- transform(mlra_acres, mlra_percent = round(mlra_percent / sum(mlra_percent) * 100, 0))

  data.l <- list(data = data, mupolygon = mupolygon, sapolygon = sapolygon)
  save(data.l, file = sample_cache)
} else load(file = sample_cache) #load sample_cache file


attach(data.l)
# mlra_acres <- data.l$mlra_acres

data$ssa_musym <- with(data, paste(areasymbol, musym))
data2 <- data
data2$ssa_musym <- "*mlra_mapnit"
data <- rbind(data2, data)
```


# Map Units
```{r mapunits}
datatable(project[!is.na(project$musym), c("projectname", "areasymbol", "musym", "muacres")])
```


# Variables

```{r variables}
a <- c("elev", "slope", "aspect", "valley", "wetness", "relief", "ppt", "temp", "ffp", "lulc")
m <- c("elevation", "slope gradient", "slope aspect", "multiresolution valley bottom index", "topographic Wetness index", "height above channel", "annual precipitation", "annual air temperature", "frost free period", "land use and land cover")
u <- c("meters", "percent", "degrees", "unitless", "unitless", "meters", "millimeters", "degrees Celsius", "days", "landcover class (e.g. Wood Wetlands)")
s <- c("30-meter USGS National Elevation Dataset (NED)", "10-meter NED", "10-meter NED", "30-meter NED", "30-meter NED", "30-meter NED", "800-meter 30-year normals (1981-2010) from PRISM Climate Dataset", "800-meter 30-year normals (1981-2010) from PRISM Climate Dataset", "1000-meter 30-year normals (1961-1990) from USFS RMRS", "2011 National Land Cover Dataset (NLCD)")

variables <- data.frame(Abbreviation = a, Measures = m, Unit = u, Source = s)

knitr::kable(variables)
```


# Map of soil polygons

```{r plot soil map units}
st <- map("state", plot = FALSE)
st_sp <- {
  map2SpatialLines(st, proj4string = CRS("+init=epsg:4326")) ->.;
  spTransform(., CRS("+init=epsg:5070"))
  }
plot(as(mupolygon, "Spatial"), axes = TRUE)
plot(as(sapolygon, "Spatial"), add = TRUE)
plot(st_sp,lwd=3, add = TRUE)

```

# Soil polygon metrics
Five number summary (min, 25th, median, 75th, max)(percentiles) and contingency table (counts)(n) 
Circularity is an estimate of SHAPE complexity (Hole and Campbell, 1975), computed as a ratio of mupolygon length / mupolygon circumference. The SHAPE complexity of a perfect circle would equal 1.

```{r soil polygon metrics, results='asis', fig.width=8}
pol <- data.frame(mupolygon)
pol$ssa_musym <- as.character(paste(pol$areasymbol, pol$musym))
pol2 <- pol
pol2$ssa_musym <- "*mlra_mapnit"
pol <- rbind(pol2, pol)

# pol.lo1 <- melt(pol, id.vars="ssa_musym", measure.vars=c("Acres", "Circularity"))
vars <- c("Acres", "Circularity")
pol.lo1 <- reshape(pol[c("ssa_musym", vars)], 
                   direction = "long",
                   timevar = "variable", times = vars,
                   v.names = "value"   , varying = vars
                   )
# pol.lo2 <- melt(pol, id.vars="ssa_musym", measure.vars=c("Acres"))
pol.lo2 <- subset(pol.lo1, variable == "Acres")

# pol.5n1 <- ddply(pol.lo2, .(ssa_musym, variable), summarize, 
#                  range = prettySummary(value, n = FALSE, signif = FALSE)
#                  )
vars <- c("ssa_musym", "variable")
pol.5n1 <- {
  by(pol.lo2, pol.lo2[vars], function(x) { data.frame(
  x[1, vars],
  range = prettySummary(x$value, n = FALSE, signif = FALSE)
  )}) ->.; 
  do.call("rbind", .) ->.;
  }
# pol.5n2 <- ddply(pol, .(ssa_musym), summarize, 
#                  nArces = round(sum(Acres), 0), 
#                  nPolygons = length(musym)
#                  )
vars <- "ssa_musym"
pol.5n2 <- {
  by(pol, pol[vars], function(x) { data.frame(
  x[1, vars, drop = FALSE],
  nAcres = round(sum(x$Acres, 0)),
  nPolygons = length(x$musym)
  )}) ->.;
  do.call("rbind", .) ->.;
  }
# pol.5n <- join(pol.5n1, pol.5n2, by = "ssa_musym")
pol.5n <- merge(pol.5n1, pol.5n2, by = "ssa_musym", all.x = TRUE)

datatable(pol.5n, caption = "Summary of musym by areasymbol")

# kable(mlra_acres, digits = 0, align = "c", caption = "Summary of Acres by MLRA")

# pol.lo1$ssa_musym <- as.factor(pol.lo1$ssa_musym)
pol.lo1$ssa_musym <- factor(pol.lo1$ssa_musym, 
                            levels = rev(sort(unique(pol.lo1$ssa_musym)))
                            )

bwplot(ssa_musym ~ value | variable, 
       data = pol.lo1, 
       scales = list(x="free"), main = "Boxplots of polygon metrics", 
       axis = axis.grid,
       as.table = TRUE
       )
# ggplot(pol.lo1, aes(x = ssa_musym, y = value)) +
#   geom_boxplot() +
#   facet_wrap(~ variable, scales = "free_x") +
#   coord_flip() +
#   ggtitle("Boxplots of polygon metrics")
```

# Contingency tables (percent) 

```{r percentage by musym, results='asis'}
## Create descriptive and graphical summary of map unit
datatable({
  round(prop.table(xtabs(~ ssa_musym + slope_classes, data = data, drop.unused.levels = TRUE),
                   margin = 1) * 100) ->.;
  as.data.frame(.) ->.;
  reshape(., direction = "wide", idvar = "ssa_musym", v.names = c("Freq"), timevar = "slope_classes"
          ) ->.;
  names(.) <- gsub("Freq.", "", names(.))
  . ->.;
  }, caption="Slope classes"
  )
datatable({
  round(prop.table(xtabs(~ ssa_musym + aspect_classes, data = data, drop.unused.levels = TRUE), 
                   margin = 1) * 100) ->.;
  as.data.frame(.) ->.;
  reshape(., direction = "wide", idvar = "ssa_musym", v.names = c("Freq"), timevar = "aspect_classes"
          ) ->.;
  names(.) <- gsub("Freq.", "", names(.))
  . ->.;
  }, caption = "Aspect classes"
  )
datatable({
  round(prop.table(xtabs(~ ssa_musym + valley_classes, data = data, drop.unused.levels = TRUE),
                   margin = 1) * 100) ->.;
  as.data.frame(.) ->.;
  reshape(., direction = "wide", idvar = "ssa_musym", v.names = c("Freq"), timevar = "valley_classes"
          ) ->.;
  names(.) <- gsub("Freq.", "", names(.))
  . ->.;
  }, caption = "Upland vs. lowland"
  )
lulc_t <- {
  xtabs(~ ssa_musym + lulc_classes, data = data, drop.unused.levels = TRUE) ->.;
  addmargins(prop.table(., margin = 1)) * 100
  }
idx <- pIndex(lulc_t, 8)
for (i in unique(idx)){
  datatable({
    round(lulc_t[, c(idx == i)]) ->.;
    as.data.frame(.) ->.;
    reshape(., direction = "wide", idvar = "ssa_musym", v.names = c("Freq"), timevar = "lulc_classes"
          ) ->.;
    names(.) <- gsub("Freq.", "", names(.))
    . ->.;
    }, caption = "Landuse and Landcover"
    )
  }
```

# Percentile breaks
Five number summary (min, 25th, median, 75th, max)(percentiles) and number of random samples (n)

```{r quantiles by musym, results="asis", warning=FALSE, fig.height=10, fig.width=8}
data.lo <- melt(data, id.vars="ssa_musym", measure.vars = c("elev", "slope", "valley", "wetness", "relief", "ppt", "temp", "ffp"))

data.5n <- ddply(data.lo, .(variable, ssa_musym), summarize,
                 range = prettySummary(value, n = FALSE, signif = FALSE)
                 )

data.c <- dcast(data.5n, ssa_musym ~ variable, value.var = 'range')
data.n <- ddply(data, .(ssa_musym), .drop=T, summarize, n = length(ssa_musym))

aspect.lo <- melt(data, id.vars="ssa_musym", measure.vars = c("aspect"))
aspect.lo$value <- circular(aspect.lo$value, template = "geographic", units="degrees", modulo="2pi")
aspect.5n <- ddply(aspect.lo, .(variable, ssa_musym), summarize,
                   range = prettySummary(value, n = FALSE, signif = FALSE)
                   )
aspect.c <- dcast(aspect.5n, ssa_musym ~ variable, value.var = 'range')

datatable(cbind(data.c[c(1:6)], data.n["n"]))
datatable(cbind(data.c[c(1, 7:9)], aspect.c["aspect"], data.n["n"]))

data.lo$ssa_musym <- factor(data.lo$ssa_musym, 
                            levels = rev(sort(unique(data.lo$ssa_musym)))
                            )

bwplot(ssa_musym ~ value | variable, 
       data = data.lo, 
       scales = list(x = "free"), main = "Boxplots of map unit properties",
       axis = axis.grid,
       as.table = TRUE, layout = c(2, 4)
       )
```
