---
title: null
output:
  html_vignette:
    mathjax: null
    jquery: null
    smart: no
    keep_md: no
    css: style.css
---


```{r setup, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(knitr, quietly=TRUE)

# package options
opts_knit$set(message=FALSE, warning=FALSE, verbose=FALSE, progress=FALSE)

# chunk options

# R session options
options(width=100, stringsAsFactors=FALSE)

## load dependencies
library(daff, quietly=TRUE)
library(soilDB, quietly=TRUE)

## load report-specific functions
source('custom.R') 

## load local configuration 
# TODO: allow entry of simple fields interactively if they are not specified?
# TODO: allow for batching from a basic report.Rmd
source('config.R')

# must pre-load selected set with relevant DMU
x <- fetchLIMS_component(project.name)

## site-level diff
s <- site(x$spc)

# remove IDs from diff
exclude.vars <- c('coiid', 'dmuiid')
site.vars <- which(! names(s) %in% exclude.vars)

# extract 'old' vs. 'new' records
a <- subset(s[, site.vars], subset=dmudesc == old.dmu)
b <- subset(s[, site.vars], subset=dmudesc == new.dmu)


## copmute and render diff
# TODO: eval optimal arguments
delta <- diff_data(a, b)
site.diff <- render_diff(delta, fragment=TRUE, view = FALSE, title = 'Site Diff', summary = TRUE)


## hz-level diff
d <- as(x$spc, 'data.frame')

hz.vars <- c('dmudesc', 'compname', horizonNames(x$spc))
exclude.vars <- c('coiid', 'chiid')
hz.vars <- hz.vars[which(! hz.vars %in% exclude.vars)]
d <- d[, hz.vars]

# split into lists
comp.order <- names(sort(tapply(s$comppct_r, s$compname, mean), decreasing = TRUE))
d$compname <- factor(d$compname, levels=comp.order)
l <- split(d, d$compname)

hz.diff <- lapply(l, function(i) {
  
  a <- subset(i, subset=dmudesc == old.dmu)
  b <- subset(i, subset=dmudesc == new.dmu)
  
  # exclude vars used for grouping
  a$compname <- NULL ; a$comppct_r <- NULL ; a$dmudesc <- NULL
  b$compname <- NULL ; b$comppct_r <- NULL ; b$dmudesc <- NULL
  
  delta <- diff_data(a, b)
  
  diff.name <- paste0('Component/Horizon Diff: ', unique(i$compname))
  # fname <- paste0('hz-diff-', unique(i$compname), '.html')
  
  res <- render_diff(delta, fragment=TRUE, view = FALSE, title = diff.name, summary = TRUE)
  return(list(title=diff.name, html=res))
  
})
```

<br>
<div style="text-align: center; border-top-style: solid; border-bottom-style: solid; border-top-width: 2px; border-bottom-width: 2px;">
<span style="font-size: 200%; font-weight: bold;">DMU Differences</span>
<span style="font-size: 100%;">
<br>
<br>
`r old.dmu`<br>vs.<br>`r new.dmu`
</span>
<br><br>
report version `r .report.version`
<br>
`r format(Sys.time(), "%Y-%m-%d")`</div>

<br>
This report provides [text differencing](https://en.wikipedia.org/wiki/Diff_utility) capabilities, applied to DMU/component records and component/horizon records queried from the selected set. This type of analyis requires marking records as *old* (`r old.dmu`) and *new* (`r new.dmu`) in the associated `config.R` file. Differences are higlighted in blue, deletions are in red, and additions are in green.


```{r, results='asis', echo=FALSE}
wrapDaffHTML(site.diff, "DMU/Component Diff")
```

```{r, results='asis', echo=FALSE}
f <- function(i) {
  wrapDaffHTML(i$html, i$title)
}

dev.null <- lapply(hz.diff, f)
```



----------------------------
This document is based on `soilDB` version `r utils::packageDescription("soilDB", field="Version")`, and `daff` version `r utils::packageDescription("daff", field="Version")`.
<br>
Report [configuration and source code are hosted on GitHub](https://github.com/ncss-tech/soilReports).



