---
title: null
output:
  html_vignette:
    mathjax: null
    jquery: null
    smart: no
    keep_md: no
    css: style.css
---


```{r setup, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(knitr, quietly=TRUE)

# chunk options
opts_knit$set(message=FALSE, warning=FALSE, verbose=FALSE, progress=FALSE)

# R session options
options(width=100, stringsAsFactors=FALSE)

## load dependencies
library(daff, quietly=TRUE)
library(soilDB, quietly=TRUE)

## load report-specific functions
source('custom.R') 

## load local configuration 
# TODO: allow entry of simple fields interactively if they are not specified?
# TODO: allow for batching from a basic report.Rmd
source('config.R')

## TODO: this would be ideal, but not quite ready
# x <- fetchLIMS_component(project.name)


## local NASIS DB setup
# load project by name
# add data for new MLRA MU via project mapunit -> mapunit table
# add data for new MLRA DMU via mapunit/correlation -> data mapunit table
x <- fetchNASIS_components()

# establish DMU vintage
x$vintage <- rep('old', times=length(x))

# flag new DMU
# could also use repdmu == 1
x$vintage[grep('MLRA', x$dmudesc)] <- 'new'
x$dmuname <- sprintf('%s\n%s', x$compname, x$vintage)

# get correlation data so we can group via new map unit name
nc <- get_component_correlation_data_from_NASIS_db(dropNotRepresentative = FALSE)
site(x) <- nc

## site-level diff
s <- site(x)

# remove IDs from diff
exclude.vars <- c('coiid', 'dmuiid')
site.vars <- which(! names(s) %in% exclude.vars)


## just in case, differences must be done by map unit
# split by muname
l <- split(s, s$muname)
site.diff <- lapply(l, function(i) {
  
  # extract 'old' vs. 'new' records
  a <- subset(i[, site.vars], subset=vintage == 'old')
  b <- subset(i[, site.vars], subset=vintage == 'new')
  
  # exclude vars used for grouping
  a$muname <- NULL ; a$areasmybol <- NULL ; a$repdmu <- NULL ; a$dmuname <- NULL ; a$muiid <- NULL ; a$vintage <- NULL
  b$muname <- NULL; b$areasmybol <- NULL ; b$repdmu <- NULL ; b$dmuname <- NULL ; b$muiid <- NULL ; b$vintage <- NULL
  
  # name for each difference
  diff.name <- unique(i$muname)
  
  ## copmute and render diff
  delta <- diff_data(a, b)
  res <- render_diff(delta, fragment=TRUE, view = FALSE, title = diff.name, summary = TRUE)
  
  return(list(title=diff.name, html=res))
})



## hz-level diff
d <- as(x, 'data.frame')

# must include map unit name for grouped comparisons
hz.vars <- c('muname', 'vintage', 'dmudesc', 'compname', horizonNames(x))
exclude.vars <- c('coiid', 'chiid')
hz.vars <- hz.vars[which(! hz.vars %in% exclude.vars)]
d <- d[, hz.vars]

# split by MU/component names
comp.order <- names(sort(tapply(s$comppct_r, s$compname, mean), decreasing = TRUE))
d$compname <- factor(d$compname, levels=comp.order)
d$muname <- factor(d$muname)
l <- split(d, list(d$muname, d$compname))

hz.diff <- lapply(l, function(i) {
  
  a <- subset(i, subset=vintage == 'old')
  b <- subset(i, subset=vintage == 'new')
  
  # exclude vars used for grouping
  a$compname <- NULL ; a$comppct_r <- NULL ; a$dmudesc <- NULL ; a$muname <- NULL ; a$vintage <- NULL
  b$compname <- NULL ; b$comppct_r <- NULL ; b$dmudesc <- NULL ; b$muname <- NULL ; b$vintage <- NULL
  
  delta <- diff_data(a, b)
  
  diff.name <- sprintf('%s/%s', unique(i$muname), unique(i$compname))
  
  res <- render_diff(delta, fragment=TRUE, view = FALSE, title = diff.name, summary = TRUE)
  return(list(title=diff.name, html=res))
  
})
```

<br>
<div style="text-align: center; border-top-style: solid; border-bottom-style: solid; border-top-width: 2px; border-bottom-width: 2px;">
<span style="font-size: 200%; font-weight: bold;">DMU Differences</span>
<span style="font-size: 100%;">
<br>
<br>
`r project.name`
</span>
<br><br>
report version `r .report.version`
<br>
`r format(Sys.time(), "%Y-%m-%d")`</div>

<br>
This report provides [text differencing](https://en.wikipedia.org/wiki/Diff_utility) capabilities, applied to DMU/component records and component/horizon records queried from the selected set. Differences are higlighted in blue, deletions are in red, and additions are in green.


```{r, results='asis', echo=FALSE}
f <- function(i) {
  wrapDaffHTML(i$html, i$title)
}

dev.null <- lapply(site.diff, f)
```

```{r, results='asis', echo=FALSE}
f <- function(i) {
  wrapDaffHTML(i$html, i$title)
}

dev.null <- lapply(hz.diff, f)
```


----------------------------
This document is based on `soilDB` version `r utils::packageDescription("soilDB", field="Version")`, and `daff` version `r utils::packageDescription("daff", field="Version")`.
<br>
Report [configuration and source code are hosted on GitHub](https://github.com/ncss-tech/soilReports).

