---
title: "EDIT Soil Features Report"
output: 
  html_document:
    toc: true
    toc_float: true
params:
  TARGET_ECOSITE_ID: "R018XI105CA"
  SELECTED_SET: false
  EXCLUDE_COMPNAME: ""
  NASIS_DSN: null
  SPATIAL_DSN: null
  INCLUDE_ADDITIONAL: false
  INCLUDE_NONREP: false
  USE_CACHE: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.retina = 3
)
# source("config.R")
source("utils.R")
```

```{r loadPackages, results='asis'}
cat(sprintf("# OVERVIEW (%s)", params$TARGET_ECOSITE_ID))
library(aqp)
library(soilDB)
library(sf)
library(mapview)
library(data.table)
.SD <- NULL; .N <- NULL; .I <- NULL; .LAST <- NULL
```

```{r nasisSummary, cache=params$USE_CACHE}
if (inherits(try(get("f"), silent = TRUE), 'try-error')) {
  f <- fetchNASIS("components",  duplicates = TRUE, fill = TRUE, SS = params$SELECTED_SET, dsn = params$NASIS_DSN)
  f <- .add_extended_data(f)
}
```

```{r subsetNASIS}
# iterate over target ecositeids
f2 <- subset(f, ecosite_id %in% params$TARGET_ECOSITE_ID)

# remove additional mapunits
if (is.null(params$INCLUDE_ADDITIONAL) || !params$INCLUDE_ADDITIONAL) {
  f2 <- subset(f2, mustatus != "additional")
}

# remove non-representative datamapunits
if (!is.null(params$INCLUDE_NONREP) || !params$INCLUDE_NONREP) {
  f2 <- subset(f2, repdmu)
}

# exclude components by name
if (!is.null(params$EXCLUDE_COMPNAME)) {
  f2 <- subset(f2, !compname %in% params$EXCLUDE_COMPNAME)
}
```

```{r cacheSpatial, cache=params$USE_CACHE}
if (is.null(params$SPATIAL_DSN) || !file.exists(params$SPATIAL_DSN)) {
  s2 <- fetchSDA_spatial(unique(f2$lmapunitiid), add.fields = c("mapunit.musym", "mapunit.muname"), chunk.size = 5)
} else {
  # TODO: local spatial data
}
```

```{r mapSpatial, fig.width=8, fig.align='center'}
SSURGO <- subset(s2, s2$mukey %in% f2$lmapunitiid)
mapview::mapview(SSURGO)
```

## Site Extent

```{r spatialSummary}
if (is.null(s2)) {
  stop("No spatial data found for: ", params$TARGET_ECOSITE_ID, call. = FALSE)
}
poly.area <- sf::st_area(s2)
s3 <- aggregate(poly.area, by = list(lmapunitiid = s2$mukey), function(x) {
  y <- sum(x, na.rm = TRUE)
  units(y) <- "acres"
  y
})
colnames(s3)[2] <- "muacres"
site(f2) <- s3
f2$compacres <- (f2$comppct_r / 100) * f2$muacres
```

## Mapunits

```{r mapunits}
mu <- unique(site(f2)[,c("mukey","musym","nationalmusym","muname","muacres")])
```

Total Mapunit Acres: `r as.character(round(sum(mu$muacres, na.rm = TRUE)))`

```{r mapunitTable}
mu$muacres <- round(mu$muacres)
mu <- mu[complete.cases(mu),]
knitr::kable(mu[order(mu$muacres, decreasing = TRUE), ],
             caption = "Mapunit List", row.names = FALSE,
             col.names = c("MUKEY", "MUSYM", "NMUSYM", "Mapunit Name", "Mapunit Acres"))
```

## Component Acres

Total Component Acres: `r as.character(round(sum(f2$compacres, na.rm = TRUE)))`

```{r}
common_components <- aggregate(f2$compacres, by = list(compname = f2$compname), function(x) round(sum(x, na.rm = TRUE)))
common_components <- merge(common_components, setNames(aggregate(f2$compacres, by = list(compname = f2$compname), length), c("compname", "n_components")),
                           by = 'compname', all.x = TRUE)
common_components$Spatial_Acres <- common_components$x
common_components$x <- NULL
knitr::kable(common_components[order(common_components$Spatial_Acres, decreasing = TRUE), ],
             caption = "Component Extent and Number of NASIS Components", row.names = FALSE,
             col.names = c("Component Name", "# Components", "Component Acres"))
```

## Component Profile Plot

```{r}
par(mar = c(0,0,0,0))
  
# simplify component names for grouping and horizon designations for thematic variable
f2$compname2 <- gsub(" taxadjunct| variant| family", "", f2$compname)
f2$genhz <- generalize.hz(f2$hzname, 
                          new = c("O", "A", "B", "C", "Cr or R"),
                          pat = c("^O", "^A", "^B", "^C", "[Rr]"))
groupedProfilePlot(f2,
                   groups = 'compname2',
                   group.name.offset = c(-5, -10),
                   print.id = FALSE,
                   name = NULL,
                   color = "genhz")
foo <- get('last_spc_plot', aqp.env)
legend(
  'bottomleft',
  legend = foo$legend.data$legend,
  ncol = 4,
  pch = 15,
  col = foo$legend.data$col,
  bty = "n"
)
foo <- dev.off()
```

# SLOPE

```{r slopeTable}
plot(density(f2$slope_r, na.rm = TRUE, from = 0, to = pmax(30, max(f2$slope_h) + 10), kernel = "rectangular"), main = "Slope Gradient (%)")
lines(density(f2$slope_l, na.rm = TRUE, from = 0, to = pmax(30, max(f2$slope_h) + 10), kernel = "rectangular"), lty = 2, col = "red")
lines(density(f2$slope_h, na.rm = TRUE, from = 0, to = pmax(30, max(f2$slope_h) + 10), kernel = "rectangular"), lty = 3, col = "blue")
abline(v = median(f2$slope_r, na.rm = TRUE))
abline(v = median(f2$slope_l, na.rm = TRUE), lty = 2, col = "red")
abline(v = median(f2$slope_h, na.rm = TRUE), lty = 3, col = "blue")
legend("topright", legend = c("Low", "RV", "High"), lty = c(2, 1, 3), col = c("RED", "BLACK", "BLUE"))
ll <- list(slope_l = f2$slope_l, 
           slope_r = f2$slope_r, 
           slope_h = f2$slope_h,
           all = c(
             f2$slope_l,
             f2$slope_r,
             f2$slope_h
           ))
knitr::kable(cbind(Variable = names(ll), do.call('rbind', lapply(ll, .quantfun))), row.names = FALSE)
```

# RUNOFF

```{r runoffTable}
f2.ro <- table(f2$runoff)
f2.ro.acres <- aggregate(f2$compacres, by = list(runoff = f2$runoff), function(x) round(sum(x, na.rm = TRUE)))
f2.ro.acres$n <- f2.ro[pmatch(f2.ro.acres$runoff, names(f2.ro))]
knitr::kable(f2.ro.acres[order(f2.ro.acres$x, decreasing = TRUE),],
             col.names = c("Local Runoff Class", "Component Acres", "# Components"), 
             row.names = FALSE)
```

# ELEVATION

```{r elevTable}
plot(density(f2$elev_r, na.rm = TRUE, from = pmin(0, min(f2$elev_l, na.rm=TRUE)), to = pmax(100, max(f2$elev_h, na.rm=TRUE)+100), kernel = "rectangular"), main = "Elevation (m)")
lines(density(f2$elev_l, na.rm = TRUE, from = pmin(0, min(f2$elev_l, na.rm=TRUE)), to = pmax(100, max(f2$elev_h, na.rm=TRUE)+100), kernel = "rectangular"), lty = 2, col = "red")
lines(density(f2$elev_h, na.rm = TRUE, from = pmin(0, min(f2$elev_l, na.rm=TRUE)), to = pmax(100, max(f2$elev_h, na.rm=TRUE)+100), kernel = "rectangular"), lty = 3, col = "blue")
abline(v = median(f2$elev_r, na.rm = TRUE))
abline(v = median(f2$elev_l, na.rm = TRUE), lty = 2, col = "red")
abline(v = median(f2$elev_h, na.rm = TRUE), lty = 3, col = "blue")
legend("topright", legend = c("Low", "RV", "High"), lty = c(2, 1, 3), col = c("RED", "BLACK", "BLUE"))
ll <- list(elev_l = f2$elev_l, 
           elev_r = f2$elev_r, 
           elev_h = f2$elev_h,
           all = c(
             f2$elev_l,
             f2$elev_r,
             f2$elev_h
           ))
elevtab <- do.call('rbind', lapply(ll, .quantfun))
knitr::kable(cbind(Variable = names(ll), round(elevtab)), row.names = FALSE)
```

## Elevation (feet)
```{r elevImperialTable}
elevtab2 <- elevtab
elevtab2[,1:9] <- elevtab2[,1:9] / 0.3048
knitr::kable( cbind(Variable = names(ll), round(elevtab2)), row.names = FALSE)
```

# FROST-FREE DAYS

```{r ffdTable}
plot(density(f2$ffd_r, na.rm = TRUE, from = 0, to = 365, kernel = "rectangular"), main = "Elevation (m)")
lines(density(f2$ffd_l, na.rm = TRUE, from = 0, to = 365, kernel = "rectangular"), lty = 2, col = "red")
lines(density(f2$ffd_h, na.rm = TRUE, from = 0, to = 365, kernel = "rectangular"), lty = 3, col = "blue")
abline(v = median(f2$ffd_r, na.rm = TRUE))
abline(v = median(f2$ffd_l, na.rm = TRUE), lty = 2, col = "red")
abline(v = median(f2$ffd_h, na.rm = TRUE), lty = 3, col = "blue")
legend("topright", legend = c("Low", "RV", "High"), lty = c(2, 1, 3), col = c("RED", "BLACK", "BLUE"))
ll <- list(ffd_l = f2$ffd_l, 
           ffd_r = f2$ffd_r, 
           ffd_h = f2$ffd_h,
           all = c(
             f2$ffd_l,
             f2$ffd_r,
             f2$ffd_h
           ))
ffdtab <- do.call('rbind', lapply(ll, .quantfun))
knitr::kable(cbind(Variable = names(ll), round(ffdtab)), 
             row.names = FALSE)
```

# PARENT MATERIAL 

## Kind

```{r pmkindTable}
f2.pm <- table(f2$pmkind)
f2.pm.acres <- aggregate(f2$compacres, by = list(pmkind = f2$pmkind), function(x) round(sum(x, na.rm = TRUE)))
f2.pm.acres$n <- f2.pm[pmatch(f2.pm.acres$pmkind, names(f2.pm))]
knitr::kable(f2.pm.acres[order(f2.pm.acres$x, decreasing = TRUE),],
             col.names = c("Parent Material Kind", "Component Acres", "# Components"), 
             row.names = FALSE)
```

## Origin

```{r pmoriginTable}
pmo <- gsub("NA & ", "", f2$pmorigin)
f2.pm <- table(pmo)
f2.pm.acres <- aggregate(f2$compacres, by = list(pmorigin = pmo), function(x) round(sum(x, na.rm = TRUE)))
f2.pm.acres$n <- f2.pm[pmatch(f2.pm.acres$pmorigin, names(f2.pm))]
knitr::kable(f2.pm.acres[order(f2.pm.acres$x, decreasing = TRUE),],
             col.names = c("Parent Material Origin", "Component Acres", "# Components"),
             row.names = FALSE)
```

# FLOODING & PONDING

```{r floodpondTable}
f2.flodcl <- table(f2$floodclass)
f2.pondcl <- table(f2$pondclass)
f2.flod.acres <- aggregate(f2$compacres, by = list(flodcl = f2$floodclass), function(x) round(sum(x, na.rm = TRUE)))
f2.pond.acres <- aggregate(f2$compacres, by = list(pondcl = f2$pondclass), function(x) round(sum(x, na.rm = TRUE)))
f2.flod.acres$n <- f2.flodcl[pmatch(f2.flod.acres$flodcl, names(f2.flodcl))]
f2.pond.acres$n <- f2.pondcl[pmatch(f2.pond.acres$pondcl, names(f2.pondcl))]

knitr::kable(f2.flod.acres[order(f2.flod.acres$x, decreasing = TRUE),],
             col.names = c("Flooding Conditions", "Component Acres", "# Components"),
             row.names = FALSE)

knitr::kable(f2.pond.acres[order(f2.pond.acres$x, decreasing = TRUE),],
             col.names = c("Ponding Conditions", "Component Acres", "# Components"),
             row.names = FALSE)
```

# DRAINAGE CLASS

```{r drainageTable}
drainage <- f2$drainagecl
f2.drainage <- table(drainage)
f2.drainage.acres <- aggregate(f2$compacres, by = list(drainagecl = drainage), function(x) round(sum(x, na.rm = TRUE)))
f2.drainage.acres$n <- f2.drainage[pmatch(f2.drainage.acres$drainagecl, names(f2.drainage))]
knitr::kable(f2.drainage.acres[order(f2.drainage.acres$x, decreasing = TRUE),],
             col.names = c("Drainage Class", "Component Acres", "# Components"),
             row.names = FALSE)
```

# SOIL DEPTH

```{r soilDepthTables}
depth.cls <- c(
  very.shallow = 25,
  shallow = 50,
  mod.deep = 102,
  deep = 152,
  very.deep = 10000
)
depth_range <- data.frame(
  l = findInterval(f2$resdept_l, depth.cls),
  rv = findInterval(f2$resdept_r, depth.cls),
  h = findInterval(f2$resdept_h, depth.cls)
) 
depth_range[] <- lapply(depth_range, function(x) {
  y <- names(depth.cls)[x + 1] 
  y[is.na(y)] <- "very.deep"
  y
})
ll <- list(resdept_l = f2$resdept_l, 
           resdept_r = f2$resdept_r, 
           resdept_h = f2$resdept_h,
           all = c(f2$resdept_l, f2$resdept_r, f2$resdept_h))
knitr::kable(cbind(Variable = names(ll), do.call('rbind', lapply(ll, .quantfun))), row.names = FALSE)
f2.depth.rv <- table(depth_range$rv, useNA = "ifany")
f2.depth.rv.acres <- aggregate(f2$compacres, by = list(rvdepth = depth_range$rv), function(x) round(sum(x, na.rm = TRUE)))
f2.depth.rv.acres$n <- f2.depth.rv[f2.depth.rv.acres$rvdepth]
knitr::kable(f2.depth.rv.acres[order(f2.depth.rv.acres$x, decreasing = TRUE),],
             col.names = c("RV Soil Depth Class", "Component Acres", "# Components"),
             row.names = FALSE)

f2.depth.l <- table(depth_range$l, useNA = "ifany")
f2.depth.l.acres <- aggregate(f2$compacres, by = list(ldepth = depth_range$l), function(x) round(sum(x, na.rm = TRUE)))
f2.depth.l.acres$n <- f2.depth.l[f2.depth.l.acres$ldepth]
knitr::kable(f2.depth.l.acres[order(f2.depth.l.acres$x, decreasing = TRUE),],
             col.names = c("Low Soil Depth Class", "Component Acres", "# Components"),
             row.names = FALSE)

f2.depth.h <- table(depth_range$h, useNA = "ifany")
f2.depth.h.acres <- aggregate(f2$compacres, by = list(hdepth = depth_range$h), function(x) round(sum(x, na.rm = TRUE)))
f2.depth.h.acres$n <- f2.depth.h[f2.depth.h.acres$hdepth]
knitr::kable(f2.depth.h.acres[order(f2.depth.h.acres$x, decreasing = TRUE),],
             col.names = c("High Soil Depth Class", "Component Acres", "# Components"),
             row.names = FALSE)
```

# PARTICLE SIZE CLASS

```{r}
f2$taxpartsize[grepl("psamm", f2$taxgrtgroup)] <- "sandy"
f2.pscs <- table(f2$taxpartsize, useNA = "ifany")
f2.pscs.acres <- aggregate(f2$compacres, by = list(taxpartsize = f2$taxpartsize), function(x) round(sum(x, na.rm = TRUE)))
f2.pscs.acres$n <- f2.pscs[f2.pscs.acres$taxpartsize]
knitr::kable(f2.pscs.acres[order(f2.pscs.acres$x, decreasing = TRUE),],
             col.names = c("Particle Size Class", "Component Acres", "# Components"),
             row.names = FALSE)
```

# MINERAL SURFACE TEXTURE

## Surface Modifier + Texture Class
```{r}
f2$mss <- getMineralSoilSurfaceDepth(f2, simplify = FALSE)$hzdepb_r
f2.first <- glom(f2, f2$mss, f2$mss + 1)[, 1]
f2.texture <- table(f2.first$texture)
f2.texture.acres <- aggregate(f2.first$compacres, by = list(texture = f2.first$texture), function(x) round(sum(x, na.rm = TRUE)))
f2.texture.acres$n <- f2.texture[f2.texture.acres$texture]
knitr::kable(f2.texture.acres[order(f2.texture.acres$x, decreasing = TRUE),],
             col.names = c("Surface Texture", "Component Acres", "# Components"),
             row.names = FALSE)
```

## Surface Texture Class
```{r}
f2.first$texcl <- gsub(".*-(.*)|(.*)", "\\1\\2", f2.first$texture)
f2.texcl.acres <- aggregate(f2.first$compacres, by = list(texcl = f2.first$texcl), function(x) round(sum(x, na.rm = TRUE)))
f2.texcl <- table(f2.first$texcl)
f2.texcl.acres$n <- f2.texcl[f2.texcl.acres$texcl]
knitr::kable(f2.texcl.acres[order(f2.texcl.acres$x, decreasing = TRUE),],
             col.names = c("Surface Texture Class", "Component Acres", "# Components"),
             row.names = FALSE)
```

## Surface Modifier
```{r}
f2.first$texmod <- gsub("(.*)-.*|.*", "\\1", f2.first$texture)
f2.first$texmod[f2.first$texmod == ""] <- "No modifier"
f2.texmod.acres <- aggregate(f2.first$compacres, by = list(texmod = f2.first$texmod), function(x) round(sum(x, na.rm = TRUE)))
f2.texmod <- table(f2.first$texmod)
f2.texmod.acres$n <- f2.texmod[f2.texmod.acres$texmod]
knitr::kable(f2.texmod.acres[order(f2.texmod.acres$x, decreasing = TRUE),],
             col.names = c("Surface Texture Modifier", "Component Acres", "# Components"),
             row.names = FALSE)
```

```{r truncateSPC}
f2$mss <- pmax(getMineralSoilSurfaceDepth(f2, simplify = FALSE)$hzdepb_r, 0, na.rm = TRUE)
f2$bdep <- pmin(f2$resdept_r, f2[, , .LAST]$hzdepb_r, na.rm = TRUE)
f2$missingDepth <- checkHzDepthLogic(f2, byhz = TRUE)$missingDepth
fsolum <- trunc(f2, f2$mss, f2$bdep)
# for truncated SPC depth subsets remove missing depths
f2v <- subsetHz(f2, !missingDepth)
f0040 <- trunc(f2v, 0, round(40 * 2.54))
f1040 <- trunc(f2v, round(10 * 2.54), round(40 * 2.54))
f0010 <- trunc(f2v, 0, round(10 * 2.54))
f0060 <- trunc(f2v, 0, round(60 * 2.54))
```

# SURFACE FRAGS <3"
```{r, results='asis'}
f0060 <- mutate_profile(f0060, 
                        surface_frags_lt3_l = surfacegravel_l,
                        surface_frags_lt3_r = surfacegravel_r,
                        surface_frags_lt3_h = surfacegravel_h)
if (sum(!is.na(f0060$surface_frags_lt3_l)) > 1 &
    sum(!is.na(f0060$surface_frags_lt3_r)) > 1 &
    sum(!is.na(f0060$surface_frags_lt3_h)) > 1) {
  plot(density(f0060$surface_frags_lt3_r, na.rm = TRUE, from = 0, kernel = "rectangular"), main = "Surface Frags <3\", Cover %")
  lines(density(f0060$surface_frags_lt3_l, na.rm = TRUE, from = 0, kernel = "rectangular"), lty = 2, col = "red")
  lines(density(f0060$surface_frags_lt3_h, na.rm = TRUE, from = 0, kernel = "rectangular"), lty = 3, col = "blue")
  abline(v = median(f0060$surface_frags_lt3_r, na.rm = TRUE))
  abline(v = median(f0060$surface_frags_lt3_l, na.rm = TRUE), lty = 2, col = "red")
  abline(v = median(f0060$surface_frags_lt3_h, na.rm = TRUE), lty = 3, col = "blue")
  legend("topright", legend = c("Low", "RV", "High"), lty = c(2, 1, 3), col = c("RED", "BLACK", "BLUE"))
}
ll <- list(surface_frags_lt3_l = f0060$surface_frags_lt3_l, 
           surface_frags_lt3_r = f0060$surface_frags_lt3_r, 
           surface_frags_lt3_h = f0060$surface_frags_lt3_h,
           all = c(
             f0060$surface_frags_lt3_l,
             f0060$surface_frags_lt3_r,
             f0060$surface_frags_lt3_h
           ))
knitr::kable(cbind(Variable = names(ll), do.call('rbind', lapply(ll, .quantfun))), row.names = FALSE)
```

# SURFACE FRAGS >=3"
```{r, results='asis'}
f0060 <- mutate_profile(f0060, 
                        surface_frags_gt3_l = surfacetotal_frags_pct_l - surfacegravel_l,
                        surface_frags_gt3_r = surfacetotal_frags_pct_r - surfacegravel_r,
                        surface_frags_gt3_h = surfacetotal_frags_pct_h - surfacegravel_h)
if (sum(!is.na(f0060$surface_frags_gt3_l)) > 1 &
    sum(!is.na(f0060$surface_frags_gt3_r)) > 1 &
    sum(!is.na(f0060$surface_frags_gt3_h)) > 1) {
  plot(density(f0060$surface_frags_gt3_r, na.rm = TRUE, from = 0, kernel = "rectangular"), main = "Surface Frags >3\", Cover %")
  lines(density(f0060$surface_frags_gt3_l, na.rm = TRUE, from = 0, kernel = "rectangular"), lty = 2, col = "red")
  lines(density(f0060$surface_frags_gt3_h, na.rm = TRUE, from = 0, kernel = "rectangular"), lty = 3, col = "blue")
  abline(v = median(f0060$surface_frags_gt3_r, na.rm = TRUE))
  abline(v = median(f0060$surface_frags_gt3_l, na.rm = TRUE), lty = 2, col = "red")
  abline(v = median(f0060$surface_frags_gt3_h, na.rm = TRUE), lty = 3, col = "blue")
  legend("topright", legend = c("Low", "RV", "High"), lty = c(2, 1, 3), col = c("RED", "BLACK", "BLUE"))
}
ll <- list(surface_frags_gt3_l = f0060$surface_frags_gt3_l, 
           surface_frags_gt3_r = f0060$surface_frags_gt3_r, 
           surface_frags_gt3_h = f0060$surface_frags_gt3_h,
           all = c(
             f0060$surface_frags_gt3_l,
             f0060$surface_frags_gt3_r,
             f0060$surface_frags_gt3_h
           ))
knitr::kable(cbind(Variable = names(ll), do.call('rbind', lapply(ll, .quantfun))), row.names = FALSE)
```

# PERMEABILITY

Most limiting, whole profile, restrictions excluded

```{r permeability, warning=FALSE}
permclasses <- c(
  Impermeable = 0,
  `Very Slow` = 0.0015,
  Slow = 0.06,
  `Moderately Slow` = 0.2,
  Moderate = 0.6,
  `Moderately Rapid` = 2.0,
  Rapid = 6.0,
  `Very Rapid` = 1000
)
fsolum$permeability <- findInterval(fsolum$ksat_l, permclasses)
if (length(fsolum) > 0) {
  fsolum <- mutate_profile(fsolum, min_permeability = as.integer(min(permeability, na.rm = TRUE)))
  # fsolum$min_permeability <- data.table::data.table(horizons(fsolum)[, c("coiid","permeability")])[, as.integer(min(as.numeric(permeability), na.rm = TRUE)), by = c("coiid")]$V1
  fsolum$min_permeability <- names(permclasses)[fsolum$min_permeability]
  fsolum.permeability <- table(fsolum$min_permeability, useNA = "ifany")
  fsolum.permeability.acres <- aggregate(fsolum$compacres, by = list(min_permeability = fsolum$min_permeability), function(x) round(sum(x, na.rm = TRUE)))
  fsolum.permeability.acres$n <- fsolum.permeability[fsolum.permeability.acres$min_permeability]
  knitr::kable(fsolum.permeability.acres[order(fsolum.permeability.acres$x, decreasing = TRUE),],
               col.names = c("Min. Permeability", "Component Acres", "# Components"),
               row.names = FALSE)
}
```

# WATER STORAGE [0 to 40 inches]

```{r aws_0to40}
f0040 <- mutate_profile(f0040, 
                        aws_l_0to40 = sum(awc_l * (hzdepb_r - hzdept_r), na.rm = TRUE) / 2.54,
                        aws_r_0to40 = sum(awc_r * (hzdepb_r - hzdept_r), na.rm = TRUE) / 2.54,
                        aws_h_0to40 = sum(awc_h * (hzdepb_r - hzdept_r), na.rm = TRUE) / 2.54)
if (length(f0040) > 2) {
  plot(density(f0040$aws_r_0to40, na.rm = TRUE, from = 0,  kernel = "rectangular"), ylim = c(0, 0.4), main = "AWS [0 to 40 in]", xlim = c(0,10))
  lines(density(f0040$aws_l_0to40, na.rm = TRUE, from = 0,  kernel = "rectangular"), lty = 2, col = "red")
  lines(density(f0040$aws_h_0to40, na.rm = TRUE, from = 0, kernel = "rectangular"), lty = 3, col = "blue")
  abline(v = median(f0040$aws_r_0to40, na.rm = TRUE))
  abline(v = median(f0040$aws_l_0to40, na.rm = TRUE), lty = 2, col = "red")
  abline(v = median(f0040$aws_h_0to40, na.rm = TRUE), lty = 3, col = "blue")
  legend("topright", legend = c("Low", "RV", "High"), lty = c(2, 1, 3), col = c("RED","BLACK","BLUE"))
}
ll <- list(aws_l_0to40 = f0040$aws_l_0to40, 
           aws_r_0to40 = f0040$aws_r_0to40, 
           aws_h_0to40 = f0040$aws_h_0to40,
           all = c(f0040$aws_l_0to40, f0040$aws_r_0to40, f0040$aws_h_0to40))
knitr::kable(cbind(Variable = names(ll), do.call('rbind', lapply(ll, .quantfun))), row.names = FALSE)
```

# REACTION 1:1 [0 to 10 inches]
<!-- pH is log-normally distributed .:. use the geometric depth weighted mean to get a value for each component -->

```{r}
f0010 <- mutate_profile(f0010, 
                        ph1to1h2o_l_0to10 = exp(weighted.mean(log(ph1to1h2o_l), hzdepb_r - hzdept_r, na.rm = TRUE)),
                        ph1to1h2o_r_0to10 = exp(weighted.mean(log(ph1to1h2o_r), hzdepb_r - hzdept_r, na.rm = TRUE)),
                        ph1to1h2o_h_0to10 = exp(weighted.mean(log(ph1to1h2o_h), hzdepb_r - hzdept_r, na.rm = TRUE)))
plot(density(f0010$ph1to1h2o_r_0to10, na.rm = TRUE, from = 3, to = 10, kernel = "rectangular"), ylim = c(0, 1.5), main = "pH [0 to 10 in]")
lines(density(f0010$ph1to1h2o_l_0to10, na.rm = TRUE, from = 3, to = 10, kernel = "rectangular"), lty = 2, col = "red")
lines(density(f0010$ph1to1h2o_h_0to10, na.rm = TRUE, from = 3, to = 10, kernel = "rectangular"), lty = 3, col = "blue")
abline(v = exp(median(log(f0010$ph1to1h2o_r_0to10), na.rm = TRUE)))
abline(v = exp(median(log(f0010$ph1to1h2o_l_0to10), na.rm = TRUE)), lty = 2, col = "red")
abline(v = exp(median(log(f0010$ph1to1h2o_h_0to10), na.rm = TRUE)), lty = 3, col = "blue")
legend("topright", legend = c("Low", "RV", "High"), lty = c(2, 1, 3), col = c("RED", "BLACK", "BLUE"))
ll <- list(ph1to1h2o_l_0to10 = log(f0010$ph1to1h2o_l_0to10), 
           ph1to1h2o_r_0to10 = log(f0010$ph1to1h2o_r_0to10), 
           ph1to1h2o_h_0to10 = log(f0010$ph1to1h2o_h_0to10),
           all = log(c(
             f0010$ph1to1h2o_l_0to10,
             f0010$ph1to1h2o_r_0to10,
             f0010$ph1to1h2o_h_0to10
           )))
knitr::kable(cbind(Variable = names(ll), do.call('rbind', lapply(ll, .quantfun, log = TRUE))), row.names = FALSE)
```

**pH classes covering 75% (or more) of RV range**

```{r, results='asis'}
prop <- prop.table(table(ph_to_rxnclass(f0010$ph1to1h2o_r_0to10)))
prop2 <- cumsum(rev(sort(prop[prop > 0])))
idx <- which(prop2 <= 0.75)
if (length(idx) == 0) idx <- 1
idx <- c(idx, max(idx) + 1)
rxn <- names(prop2[idx])
cat("   ")
cat(paste0("_", na.omit(rxn), "_"), sep = ", ")
cat("\n")
knitr::kable(rxnclass_to_ph(rxn), row.names = FALSE)
```

**(Half) pH classes covering 75% (or more) of RV range**

```{r, results='asis'}
prop <- prop.table(table(ph_to_rxnclass(f0010$ph1to1h2o_r_0to10, halfclass = TRUE)))
prop2 <- cumsum(rev(sort(prop[prop > 0])))
idx <- which(prop2 <= 0.75)
if (length(idx) == 0) idx <- 1
idx <- c(idx, max(idx) + 1)
rxn <- names(prop2[idx])
cat("   ")
cat(paste0("_", na.omit(rxn), "_"), sep = ", ")
cat("\n")
knitr::kable(rxnclass_to_ph(rxn), row.names = FALSE)
```

**Component-specific pH Low and High Class Limits**

```{r, results='asis'}
ll <- lapply(do.call('rbind', lapply(1:length(f0010), function(i) {
  rxnclass_to_ph(ph_to_rxnclass(c(f0010$ph1to1h2o_l_0to10[i], 
                                  f0010$ph1to1h2o_r_0to10[i], 
                                  f0010$ph1to1h2o_h_0to10[i]), halfclass = TRUE))
})), function(x) c(quantile(x, probs = c(0, 0.01, 0.05, 0.10, 0.5, 0.90, 0.95,  0.99, 1)), n=length(f0010)))
d <- cbind(Variable = names(ll), do.call('rbind', lapply(ll, t)))
d[, 2:ncol(d)] <- apply(d[, 2:ncol(d)], 2, function(x) round(as.numeric(x), 1))
knitr::kable(d, row.names = FALSE)
```


# REACTION 1:1 [10 to 40 inches]

<!-- pH is log-normally distributed .:. use the geometric depth weighted mean to get a value for each component -->

```{r}
f1040 <- mutate_profile(f1040, 
                        ph1to1h2o_l_10to40 = exp(weighted.mean(log(ph1to1h2o_l), hzdepb_r - hzdept_r, na.rm = TRUE)),
                        ph1to1h2o_r_10to40 = exp(weighted.mean(log(ph1to1h2o_r), hzdepb_r - hzdept_r, na.rm = TRUE)),
                        ph1to1h2o_h_10to40 = exp(weighted.mean(log(ph1to1h2o_h), hzdepb_r - hzdept_r, na.rm = TRUE)))
if (sum(!is.na(f1040$ph1to1h2o_r_10to40)) > 1 &
    sum(!is.na(f1040$ph1to1h2o_l_10to40)) > 1 &
    sum(!is.na(f1040$ph1to1h2o_h_10to40)) > 1) {
  plot(density(f1040$ph1to1h2o_r_10to40, na.rm = TRUE, from = 3, to = 10, kernel = "rectangular"), ylim = c(0, 1.5), main = "pH [0 to 10 in]")
  lines(density(f1040$ph1to1h2o_l_10to40, na.rm = TRUE, from = 3, to = 10, kernel = "rectangular"), lty = 2, col = "red")
  lines(density(f1040$ph1to1h2o_h_10to40, na.rm = TRUE, from = 3, to = 10, kernel = "rectangular"), lty = 3, col = "blue")
  abline(v = exp(median(log(f1040$ph1to1h2o_r_10to40), na.rm = TRUE)))
  abline(v = exp(median(log(f1040$ph1to1h2o_l_10to40), na.rm = TRUE)), lty = 2, col = "red")
  abline(v = exp(median(log(f1040$ph1to1h2o_h_10to40), na.rm = TRUE)), lty = 3, col = "blue")
  legend("topright", legend = c("Low", "RV", "High"), lty = c(2, 1, 3), col = c("RED", "BLACK", "BLUE"))
}
ll <- list(ph1to1h2o_l_10to40 = log(f1040$ph1to1h2o_l_10to40), 
           ph1to1h2o_r_10to40 = log(f1040$ph1to1h2o_r_10to40), 
           ph1to1h2o_h_10to40 = log(f1040$ph1to1h2o_h_10to40),
           all = log(c(
             f1040$ph1to1h2o_l_10to40,
             f1040$ph1to1h2o_r_10to40,
             f1040$ph1to1h2o_h_10to40
           )))
knitr::kable(cbind(Variable = names(ll), do.call('rbind', lapply(ll, .quantfun, log = TRUE))), row.names = FALSE)
```

**pH classes covering 75% (or more) of RV range**

```{r, results='asis'}
prop <- prop.table(table(ph_to_rxnclass(f1040$ph1to1h2o_r_10to40)))
prop2 <- cumsum(rev(sort(prop[prop > 0])))
idx <- which(prop2 <= 0.75)
if (length(idx) == 0) idx <- 1
idx <- c(idx, max(idx) + 1)
rxn <- names(prop2[idx])
cat("   ")
cat(paste0("_", na.omit(rxn), "_"), sep = ", ")
cat("\n")
knitr::kable(rxnclass_to_ph(rxn), row.names = FALSE)
```

**(Half) pH classes covering 75% (or more) of RV range**

```{r, results='asis'}
prop <- prop.table(table(ph_to_rxnclass(f1040$ph1to1h2o_r_10to40, halfclass = TRUE)))
prop2 <- cumsum(rev(sort(prop[prop > 0])))
idx <- which(prop2 <= 0.75)
if (length(idx) == 0) idx <- 1
idx <- c(idx, max(idx) + 1)
rxn <- names(prop2[idx])
cat("   ")
cat(paste0("_", na.omit(rxn), "_"), sep = ", ")
cat("\n")
knitr::kable(rxnclass_to_ph(rxn), row.names = FALSE)
```

**Component-specific pH Low and High Class Limits**

```{r, results='asis'}
ll <- lapply(do.call('rbind', lapply(1:length(f1040), function(i) {
  rxnclass_to_ph(ph_to_rxnclass(c(f1040$ph1to1h2o_l_10to40[i], 
                                  f1040$ph1to1h2o_r_10to40[i], 
                                  f1040$ph1to1h2o_h_10to40[i]), halfclass = TRUE))
})), function(x) c(quantile(x, probs = c(0, 0.01, 0.05, 0.10, 0.5, 0.90, 0.95,  0.99, 1), na.rm = TRUE), n = length(f1040)))
d <- cbind(Variable = names(ll), do.call('rbind', lapply(ll, t)))
d[, 2:ncol(d)] <- apply(d[, 2:ncol(d)], 2, function(x) round(as.numeric(x), 1))
knitr::kable(d, row.names = FALSE)
```

# SUBSURFACE FRAGS <3"
```{r, results='asis'}
f0060 <- mutate_profile(f0060, 
                        subsurface_frags_lt3_l = weighted.mean(lt3gt2mmvol_l, hzdepb_r - hzdept_r, na.rm = TRUE),
                        subsurface_frags_lt3_r = weighted.mean(lt3gt2mmvol_r, hzdepb_r - hzdept_r, na.rm = TRUE),
                        subsurface_frags_lt3_h = weighted.mean(lt3gt2mmvol_h, hzdepb_r - hzdept_r, na.rm = TRUE))
plot(density(f0060$subsurface_frags_lt3_r, na.rm = TRUE, from = 0, kernel = "rectangular"), main = "Subsurface Frags <3\" [0 to 60 in], Volume %")
lines(density(f0060$subsurface_frags_lt3_l, na.rm = TRUE, from = 0, kernel = "rectangular"), lty = 2, col = "red")
lines(density(f0060$subsurface_frags_lt3_h, na.rm = TRUE, from = 0, kernel = "rectangular"), lty = 3, col = "blue")
abline(v = median(f0060$subsurface_frags_lt3_r, na.rm = TRUE))
abline(v = median(f0060$subsurface_frags_lt3_l, na.rm = TRUE), lty = 2, col = "red")
abline(v = median(f0060$subsurface_frags_lt3_h, na.rm = TRUE), lty = 3, col = "blue")
legend("topright", legend = c("Low", "RV", "High"), lty = c(2, 1, 3), col = c("RED", "BLACK", "BLUE"))
ll <- list(subsurface_frags_lt3_l = f0060$subsurface_frags_lt3_l, 
           subsurface_frags_lt3_r = f0060$subsurface_frags_lt3_r, 
           subsurface_frags_lt3_h = f0060$subsurface_frags_lt3_h,
           all = c(
             f0060$subsurface_frags_lt3_l,
             f0060$subsurface_frags_lt3_r,
             f0060$subsurface_frags_lt3_h
           ))
knitr::kable(cbind(Variable = names(ll), do.call('rbind', lapply(ll, .quantfun))), row.names = FALSE)
```

# SUBSURFACE FRAGS >=3"
```{r, results='asis'}
f0060 <- mutate_profile(f0060, 
                        subsurface_frags_gt3_l = weighted.mean(gt3vol_l, hzdepb_r - hzdept_r, na.rm = TRUE),
                        subsurface_frags_gt3_r = weighted.mean(gt3vol_r, hzdepb_r - hzdept_r, na.rm = TRUE),
                        subsurface_frags_gt3_h = weighted.mean(gt3vol_h, hzdepb_r - hzdept_r, na.rm = TRUE))
plot(density(f0060$subsurface_frags_gt3_r, na.rm = TRUE, from = 0, kernel = "rectangular"), main = "Subsurface Frags >3\" [0 to 60 in], Volume %")
lines(density(f0060$subsurface_frags_gt3_l, na.rm = TRUE, from = 0, kernel = "rectangular"), lty = 2, col = "red")
lines(density(f0060$subsurface_frags_gt3_h, na.rm = TRUE, from = 0, kernel = "rectangular"), lty = 3, col = "blue")
abline(v = median(f0060$subsurface_frags_gt3_r, na.rm = TRUE))
abline(v = median(f0060$subsurface_frags_gt3_l, na.rm = TRUE), lty = 2, col = "red")
abline(v = median(f0060$subsurface_frags_gt3_h, na.rm = TRUE), lty = 3, col = "blue")
legend("topright", legend = c("Low", "RV", "High"), lty = c(2, 1, 3), col = c("RED", "BLACK", "BLUE"))
ll <- list(subsurface_frags_gt3_l = f0060$subsurface_frags_gt3_l, 
           subsurface_frags_gt3_r = f0060$subsurface_frags_gt3_r, 
           subsurface_frags_gt3_h = f0060$subsurface_frags_gt3_h,
           all = c(
             f0060$subsurface_frags_gt3_l,
             f0060$subsurface_frags_gt3_r,
             f0060$subsurface_frags_gt3_h
           ))
knitr::kable(cbind(Variable = names(ll), do.call('rbind', lapply(ll, .quantfun))), row.names = FALSE)
```

