---
title: null
output:
  html_vignette:
    mathjax: null
    jquery: null
    smart: no
    keep_md: no
    css: style.css
params:
  projectname: !r as.character("MLRA 17 - Arbuckle horizon standardization in 2-CHI SSA updated in SDJR")
---



```{r setup, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(knitr, quietly=TRUE)

# chunk options
opts_chunk$set(message=FALSE, warning=FALSE, background='#F7F7F7', fig.align='center', fig.retina=2, dev='png', tidy=FALSE, verbose=FALSE, progress=FALSE)

# R session options
options(width=100, stringsAsFactors=FALSE)

## load dependencies
library(soilDB, quietly=TRUE)
library(latticeExtra, quietly=TRUE)
library(wesanderson, quietly=TRUE)
library(plyr, quietly=TRUE)

## load report-specific functions
source('custom.R') 

## load local configuration 
project.name <- params$projectname

## local NASIS DB setup
# load project by name
# add data for new MLRA MU via project mapunit -> mapunit table
# add data for new MLRA DMU via mapunit/correlation -> data mapunit table

## load relevant stuff from local NASIS
# mu text + mu names / symbols
mutext <- get_mutext_from_NASIS_db()

# DMU / components as SPC
x <- fetchNASIS_components()

# MU / correlation
nc <- get_component_correlation_data_from_NASIS_db(dropNotRepresentative = FALSE)

## post-processing
# TODO: better to use mutype?
# flag vintage via pattern matching
x$vintage <- rep('old', times=length(x))

# flag new DMU
# could also use repdmu == 1
x$vintage[grep('MLRA', x$dmudesc)] <- 'new'
x$dmuname <- sprintf('%s\n%s', x$compname, x$vintage)

# join MU / correlation into DMU / component @site
site(x) <- nc


## establish some reasonable figure widths in inches
profile.fig.width <- 1 + (length(x) * 0.85)

## TODO: base this on user-supplied properties
slab.fig.width <- 12

## establish some reasonable figure heights
mu.prop.panels.height <- 1 + (length(unique(x$nationalmusym)) * 2.5)

## TODO user-specified horizon properties as parameter

```

<br>
<div style="text-align: center; border-top-style: solid; border-bottom-style: solid; border-top-width: 2px; border-bottom-width: 2px;">
<span style="font-size: 200%; font-weight: bold;">QA Summary</span>
<span style="font-size: 100%;">
<br>
<br>
`r project.name`
</span>
<br><br>
report version `r .report.version`
<br>
`r format(Sys.time(), "%Y-%m-%d")`</div>

<br>
This report requires setting up a selected set in your local NASIS database:

   * load project by name
   * add data associated with MLRA map units into *mapunit table*
   * add data associated with old and new DMU into *data mapunit table*


## MLRA Map Unit Text Notes

```{r echo=FALSE, results='asis'}

# just the MLRA map units
mutext <- mutext[which(mutext$mutype == 'mlra map unit'), ]

# iterate over rows and output nicely formatted text
for(i in 1:nrow(mutext)) {
  m <- mutext[i, ]
  
  txt <- sprintf("### %s: %s\n<div id='mu_textnote'>%s</div>\n\n", m$nationalmusym, m$muname, m$textentry)
  
  cat(txt)
}

```

## Profile Sketches
Components associated with MLRA map units are marked as "new".
```{r echo=FALSE, results='hide', fig.width=profile.fig.width, fig.height=6}

par(mar=c(1,1,4,1))
groupedProfilePlot(x, groups = 'nationalmusym', label='dmuname', group.name.cex = 0.85, group.name.offset = c(-16, -18), color='claytotal_r', col.label='Clay Content (%)', col.legend.cex=0.85)

groupedProfilePlot(x, groups = 'nationalmusym', label='dmuname', group.name.cex = 0.85, group.name.offset = c(-16, -18), color='sandtotal_r', col.label='Sand Content (%)', col.legend.cex=0.85)

groupedProfilePlot(x, groups = 'nationalmusym', label='dmuname', group.name.cex = 0.85, group.name.offset = c(-16, -18), color='fragvoltot_r', col.label='Total Fragment Volumne (%)', col.legend.cex=0.85)

groupedProfilePlot(x, groups = 'nationalmusym', label='dmuname', group.name.cex = 0.85, group.name.offset = c(-16, -18), color='texture', col.label='Texture', col.legend.cex=0.75)

groupedProfilePlot(x, groups = 'nationalmusym', label='dmuname', group.name.cex = 0.85, group.name.offset = c(-16, -18), color='ec_r', col.label='EC (mmhos/cm)', col.legend.cex=0.85)

groupedProfilePlot(x, groups = 'nationalmusym', label='dmuname', group.name.cex = 0.85, group.name.offset = c(-16, -18), color='cec7_r', col.label='CEC @ pH 7 (cmol[+] / kg)', col.legend.cex=0.85)

groupedProfilePlot(x, groups = 'nationalmusym', label='dmuname', group.name.cex = 0.85, group.name.offset = c(-16, -18), color='sumbases_r', col.label='Base Saturation (%)', col.legend.cex=0.85)


groupedProfilePlot(x, groups = 'nationalmusym', label='dmuname', group.name.cex = 0.85, group.name.offset = c(-16, -18), color='lep_r', col.label='LEP', col.legend.cex=0.85)


groupedProfilePlot(x, groups = 'nationalmusym', label='dmuname', group.name.cex = 0.85, group.name.offset = c(-16, -18), color='ph1to1h2o_r', col.label='pH 1:1 H2O', col.legend.cex=0.85)


groupedProfilePlot(x, groups = 'nationalmusym', label='dmuname', group.name.cex = 0.85, group.name.offset = c(-16, -18), color='sar_r', col.label='SAR (%)', col.legend.cex=0.85)

groupedProfilePlot(x, groups = 'nationalmusym', label='dmuname', group.name.cex = 0.85, group.name.offset = c(-16, -18), color='awc_r', col.label='AWC (cm / cm)', col.legend.cex=0.85)


groupedProfilePlot(x, groups = 'nationalmusym', label='dmuname', group.name.cex = 0.85, group.name.offset = c(-16, -18), color='ksat_r', col.label='Ksat (um/s)', col.legend.cex=0.85)


groupedProfilePlot(x, groups = 'nationalmusym', label='dmuname', group.name.cex = 0.85, group.name.offset = c(-16, -18), color='om_r', col.label='Organic Matter (%)', col.legend.cex=0.85)


groupedProfilePlot(x, groups = 'nationalmusym', label='dmuname', group.name.cex = 0.85, group.name.offset = c(-16, -18), color='dbthirdbar_r', col.label='Db 1/3 Bar (g / cc)', col.legend.cex=0.85)


```



## RV Summaries by Depth
Median component RV along 1cm slices, grouped by component name and mapunit type ("new" vs. "old").
```{r echo=FALSE, results='hide', fig.width=slab.fig.width, fig.height=6, out.width="100%"}
# quantiles over depth slices as grouped by dmuname/vintage labels
a <- slab(x, dmuname ~ claytotal_r + cec7_r + awc_r + om_r + ph1to1h2o_r, slab.fun = aqp:::.slab.fun.numeric.fast)
a$dmuname <- factor(a$dmuname)

# better plot style
cols <- brewer.pal(n=8, 'Paired')[c(2,1,4,3,6,5,8,7,10,9)]
tps <- list(superpose.line=list(lwd=2, col=cols))

xyplot(
  top ~ p.q50 | variable, data=a, groups=dmuname,
  lower=a$p.q5, upper=a$p.q95, sync.colors=TRUE, alpha=0.5,
  ylim=c(160,-5), layout=c(5,1), scales=list(x=list(relation='free')),
  xlab='', ylab='',
  par.settings=tps,
  strip = strip.custom(bg=grey(0.85)),
  panel=panel.depth_function, 
  prepanel=prepanel.depth_function,
  auto.key=list(columns=length(levels(a$dmuname)), lines=TRUE, points=FALSE)
)
```

Median component RV along 1cm slices, grouped by component name and mapunit type ("new" vs. "old"), and split into panels by national map unit symbol.
```{r echo=FALSE, results='hide', fig.width=slab.fig.width, fig.height=mu.prop.panels.height, out.width="100%"}
# iterate over MU
mu <- unique(x$nationalmusym)
ll <- list()
for(i in mu) {
  
  # subset to current MU
  # quantiles over depth slices as grouped by dmuname/vintage labels
  # just in case there are multiple "old" / "new" groups, use quantiles vs. identity
  a <- slab(x[which(x$nationalmusym == i), ], dmuname ~ claytotal_r + cec7_r + awc_r + om_r + ph1to1h2o_r, slab.fun = aqp:::.slab.fun.numeric.fast)
  
  # convert to factor for plotting
  a$dmuname <- factor(a$dmuname)
  
  # add to indexed list
  ll[[i]] <- a
}

# convert back to DF
aa <- ldply(ll)

# plot styling
cols <- brewer.pal(n=8, 'Paired')[c(2,1,4,3,6,5,8,7,10,9)]
tps <- list(superpose.line=list(lwd=2, col=cols))

p <- xyplot(
  top ~ p.q50 | variable + .id, data=aa, groups=dmuname,
  ylim=c(200,-5), 
  scales=list(x=list(relation='free'), alternating=3),
  par.settings=tps,
  panel=panel.depth_function, 
  prepanel=prepanel.depth_function,
  auto.key=list(columns=length(levels(a$dmuname)), lines=TRUE, points=FALSE)
)

# move MU groups to outerstrip
p <- useOuterStrips(p, strip = strip.custom(bg=grey(0.85)), strip.left = strip.custom(bg=grey(0.75)))

# fix axis labels and print
update(p, xlab='', ylab='')
```


----------------------------
This document is based on `soilDB` version `r utils::packageDescription("soilDB", field="Version")` and `aqp` version `r utils::packageDescription("aqp", field="Version")`.
<br>
Report [configuration and source code are hosted on GitHub](https://github.com/ncss-tech/soilReports).

