---
title: null
output:
  html_vignette:
    mathjax: null
    jquery: null
    smart: no
    keep_md: no
    css: style.css
---

```{r setup, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(knitr, quietly=TRUE)

# chunk options
opts_chunk$set(message=FALSE, warning=FALSE, background='#F7F7F7', fig.align='center', fig.retina=2, dev='png', tidy=FALSE, verbose=FALSE, progress=FALSE)

# R session options
options(width=100, stringsAsFactors=FALSE)

## load dependencies
library(soilDB, quietly=TRUE)
library(sharpshootR, quietly=TRUE)
library(latticeExtra, quietly=TRUE)
library(plyr, quietly=TRUE)
library(ggplot2, quietly=TRUE)

## load report-specific functions
source('custom.R') 

## load relevant stuff from local NASIS

# project metadata from local NASIS
# defined in custom.R
project.metadata <- get_project_meta()

# mu text + mu names / symbols
mutext <- get_mutext_from_NASIS_db()

## TODO: should we bring in those component that are missing hz data?
## probably not at this point, as profile sketchs and aggregation will be messed up
# DMU / components as SPC
x <- fetchNASIS_components()

# MU / correlation
nc <- get_component_correlation_data_from_NASIS_db(dropNotRepresentative = FALSE)

## additional stuff via WWW reports

## TODO: check for "old" DMU not linked to "new" MLRA MU
## look for NA in mu.set

# MU affected
# must supply a FY, default to current year (?)
mu.set <- get_projectmapunit2_from_NASISWebReport(mlrassoarea = '%', fiscalyear = format(Sys.Date(), "%Y"), projectname = project.metadata$projectname)

# flag records that should be removed:
mu.set$drop <- FALSE

# archived SSA
idx <- grep('arch', mu.set$areasymbol, ignore.case = TRUE)
mu.set$drop[idx] <- TRUE

# mustatus of additional / provisional
idx <- mu.set$mustatus %in% c('Provisional', 'Additional')
mu.set$drop[idx] <- TRUE

## Note: this could indicate and old DMU not linked to new MU
# mustatus of NA (except MLRA Map Unit)
idx <- which(is.na(mu.set$mustatus) & is.na(mu.set$mutype))
mu.set$drop[idx] <- TRUE

## OSD data for dendrogram
series.set <- unique(get_component_data_from_NASIS_db()$compname)
spc <- fetchOSD(series.set)
n.spc <- length(spc)

# this figure is only possible if there are more than 1 series
do.osd.dend <- n.spc > 1


## post-processing
# TODO: better to use mutype?
# flag vintage via pattern matching
x$vintage <- rep('old', times=length(x))

# flag new DMU
# could also use repdmu == 1
x$vintage[grep('MLRA', x$dmudesc)] <- 'new'
x$dmuname <- sprintf('%s\n%s', x$compname, x$vintage)

# abbreviated dmudesc, MLRA DMU have a very long name
x$dmudesc_short <- ifelse(grepl('mlra', x$dmudesc, ignore.case = TRUE), '*NEW*', x$dmudesc)

# join MU / correlation into DMU / component @site
site(x) <- nc

# still working on this...
# group name for profile plots
# x$groupname <- sprintf("%s\n%s", x$nationalmusym, x$dmudesc_short)
x$groupname <- sprintf("%s\n%s", x$nationalmusym, x$compname)

## estimate base saturation at pH 7
x$base.sat <- pmin(100, (x$sumbases_r / x$cec7_r) * 100)



## TODO: this may include some false positives for dropping: DMU linked to multiple MU

## keep only those data associated with correlated MU in active SSA
test.1 <- x$dmuiid %in% mu.set$dmuiid[which(! mu.set$drop)]
## OR vintage marked as "new" --> these are the new, MLRA DMU
test.2 <- x$vintage == 'new'

idx <- which( test.1 | test.2)
x <- x[idx, ]



## TODO: this should be split of MLRA MU symbols
## component month summaries
# load comonth
cm <- get_comonth_from_NASIS_db(fill = TRUE)
# combine DMU/component names
cm <- join(site(x), cm, by='coiid')
# re-level component names based on mean comppct
cm.compname.level.data <- sort(round(tapply(cm$comppct, cm$compname, mean)), decreasing = TRUE)
cm$compname <- factor(cm$compname, 
                      levels=names(cm.compname.level.data), 
                      labels = sprintf("%s (%s%%)", names(cm.compname.level.data), cm.compname.level.data)
                      )

## TODO: maybe add this
# # drop component names which aren't present in the MLRA MU
# missing.test <- tapply(cm$dmudesc_short == '*NEW*', cm$compname, which)
# idx <- which(sapply(missing.test, length) == 0)


## establish some reasonable figure widths in inches

# profile figures
profile.fig.width <- 1.5 + (length(x) * 0.85)

# OSD dendrogram for all components that are named series
osd.fig.width <- 5 + (n.spc * 0.5)

# slab figures
# TODO: base this on user-supplied properties
slab.fig.width <- 12

## establish some reasonable figure heights
mu.prop.panels.height <- 1 + (length(unique(x$nationalmusym)) * 2.5)


## TODO user-specified horizon properties as parameter

## TODO / ideas
##  
##  site-level comparisons in aggregate
##  table(x$dmuname, x$taxsubgrp)
##
## link to SEE or mapview of series extents
## http://ncss-tech.github.io/stats_for_soil_survey/chapters/2_data/2b_spatial_data.html
##
## table of affected components
##
## graphical breakdown of Legend -> MU -> DMU -> component by mutype
##


```

<br>
<div style="text-align: center; border-top-style: solid; border-bottom-style: solid; border-top-width: 2px; border-bottom-width: 2px;">
<span style="font-size: 200%; font-weight: bold;">QA Summary</span>
<span style="font-size: 100%;">
<br>
<br>
`r project.metadata$projectname`
</span>
<br><br>
report version `r .report.version`
<br>
`r format(Sys.time(), "%Y-%m-%d %H:%M")`</div>

<br>
This report requires setting up a selected set in your local NASIS database:

   * load project by name
   * add data associated with MLRA map units into *mapunit table*
   * add data associated with old and new DMU into *data mapunit table*


## Project Description

```{r echo=FALSE, results='asis'}
txt <- sprintf("<div id='mu_textnote'>%s</div>\n\n", project.metadata$projectdesc)
cat(txt)
```

## MLRA Map Unit Text Notes

```{r echo=FALSE, results='asis'}
# just the MLRA map units
mutext <- mutext[which(mutext$mutype == 'mlra map unit'), ]

# split by national musym
mutext.list <- split(mutext, f = mutext$nationalmusym)

# iterate over national musym
l_ply(mutext.list, function(text.notes) {
  
  # start section
  cat(sprintf("### %s: %s\n", unique(text.notes$nationalmusym), unique(text.notes$muname)))
  
  # split by text note type
  notes <- split(text.notes, text.notes$mapunittextkind)
  
  # iterate over note type
  l_ply(notes, function(this.note) {
    
    # in the case of multiple rows
    for(i in 1:nrow(this.note)) {
      txt <- sprintf("#### %s\n<div id='mu_textnote'>%s</div>\n", this.note$mapunittextkind[i], this.note$textentry[i])
      cat(txt)
    }
    
    
  })
  
})

```

## Affected Map Units
```{r echo=FALSE}
## HTML formatting isn't quite right when using html_vignette
mu.set[order(mu.set$drop, mu.set$areasymbol), c('areasymbol', 'musym', 'nationalmusym', 'mutype', 'mustatus', 'drop')]
```

## Component Breakdown
```{r echo=FALSE}
site(x)[, c('dmudesc', 'vintage', 'compname', 'comppct_r')]
```


```{r echo=FALSE, results='hide', fig.width=osd.fig.width, fig.height=6.5, eval=do.osd.dend}
SoilTaxonomyDendrogram(spc, cex.taxon.labels=0.8)
```



## Profile Sketches
Groups are formed by the new national map unit symbol / component name. Profiles associated with the "old" DMU are labeled with the parent `dmudesc`. Profiles associated with  the "new" MLRA DMU are labeled as `*NEW*`.
```{r echo=FALSE, results='hide', fig.width=profile.fig.width, fig.height=6.5}

# setup a reasonable group name offset
n.groups <- length(unique(x$groupname))
if(n.groups > 1) {
  # use an alternating group name offset
  # this is usually a good strategy
  gno <- c(-12, -14)
} else {
  # single group, use a single offset
  gno <- -13
}

par(mar=c(0.25, 0.5, 4, 1))
groupedProfilePlot(x, groups = 'groupname', label='dmudesc_short', group.name.cex = 0.75, group.name.offset = gno, color='claytotal_r', col.label='Clay Content (%)', col.legend.cex=0.85)

groupedProfilePlot(x, groups = 'groupname', label='dmudesc_short', group.name.cex = 0.75, group.name.offset = gno, color='sandtotal_r', col.label='Sand Content (%)', col.legend.cex=0.85)

groupedProfilePlot(x, groups = 'groupname', label='dmudesc_short', group.name.cex = 0.75, group.name.offset = gno, color='fragvoltot_r', col.label='Total Fragment Volumne (%)', col.legend.cex=0.85)

groupedProfilePlot(x, groups = 'groupname', label='dmudesc_short', group.name.cex = 0.75, group.name.offset = gno, color='texture', col.label='Texture', col.legend.cex=0.75)



groupedProfilePlot(x, groups = 'groupname', label='dmudesc_short', group.name.cex = 0.75, group.name.offset = gno, color='lep_r', col.label='LEP', col.legend.cex=0.85)

groupedProfilePlot(x, groups = 'groupname', label='dmudesc_short', group.name.cex = 0.75, group.name.offset = gno, color='cec7_r', col.label='CEC @ pH 7 (cmol[+] / kg)', col.legend.cex=0.85)

groupedProfilePlot(x, groups = 'groupname', label='dmudesc_short', group.name.cex = 0.75, group.name.offset = gno, color='sumbases_r', col.label='Sum of Bases (cmol[+] / kg)', col.legend.cex=0.85)

groupedProfilePlot(x, groups = 'groupname', label='dmudesc_short', group.name.cex = 0.75, group.name.offset = gno, color='base.sat', col.label='Base Saturation @ pH 7 (%)', col.legend.cex=0.85)



groupedProfilePlot(x, groups = 'groupname', label='dmudesc_short', group.name.cex = 0.75, group.name.offset = gno, color='ph1to1h2o_r', col.label='pH 1:1 H2O', col.legend.cex=0.85)

groupedProfilePlot(x, groups = 'groupname', label='dmudesc_short', group.name.cex = 0.75, group.name.offset = gno, color='ec_r', col.label='EC (mmhos/cm)', col.legend.cex=0.85)

groupedProfilePlot(x, groups = 'groupname', label='dmudesc_short', group.name.cex = 0.75, group.name.offset = gno, color='sar_r', col.label='SAR (%)', col.legend.cex=0.85)

groupedProfilePlot(x, groups = 'groupname', label='dmudesc_short', group.name.cex = 0.75, group.name.offset = gno, color='caco3_r', col.label='CaCO3 (%)', col.legend.cex=0.85)


groupedProfilePlot(x, groups = 'groupname', label='dmudesc_short', group.name.cex = 0.75, group.name.offset = gno, color='awc_r', col.label='AWC (cm / cm)', col.legend.cex=0.85)

groupedProfilePlot(x, groups = 'groupname', label='dmudesc_short', group.name.cex = 0.75, group.name.offset = gno, color='ksat_r', col.label='Ksat (um/s)', col.legend.cex=0.85)


groupedProfilePlot(x, groups = 'groupname', label='dmudesc_short', group.name.cex = 0.75, group.name.offset = gno, color='om_r', col.label='Organic Matter (%)', col.legend.cex=0.85)

groupedProfilePlot(x, groups = 'groupname', label='dmudesc_short', group.name.cex = 0.75, group.name.offset = gno, color='dbthirdbar_r', col.label='Db 1/3 Bar (g / cc)', col.legend.cex=0.85)


```

## Component Month Summaries

```{r echo=FALSE, results='hide', fig.width=10, fig.height=8}
ggplot(cm, aes(month, dmudesc_short, flodfreqcl)) + geom_tile(aes(fill = flodfreqcl), color='white', lwd=1) + scale_fill_brewer(palette = "Spectral", drop=FALSE, na.value='grey80', name='Flooding Frequency') + scale_x_discrete(drop=FALSE) + theme_minimal() + xlab('') + ylab('') + facet_wrap(vars(groupname)) + theme(legend.position="bottom")

ggplot(cm, aes(month, dmudesc_short, floddurcl)) + geom_tile(aes(fill = floddurcl), color='white', lwd=1) + scale_fill_brewer(palette = "Spectral", drop=FALSE, na.value='grey80', name='Flooding Duration') + scale_x_discrete(drop=FALSE) + theme_minimal() + xlab('') + ylab('') + facet_wrap(vars(groupname)) + theme(legend.position="bottom")


ggplot(cm, aes(month, dmudesc_short, pondfreqcl)) + geom_tile(aes(fill = pondfreqcl), color='white', lwd=1) + scale_fill_brewer(palette = "Spectral", drop=FALSE, na.value='grey80', name='Ponding  Frequency') + scale_x_discrete(drop=FALSE) + theme_minimal() + xlab('') + ylab('') + facet_wrap(vars(groupname)) + theme(legend.position="bottom")

ggplot(cm, aes(month, dmudesc_short, ponddurcl)) + geom_tile(aes(fill = ponddurcl), color='white', lwd=1) + scale_fill_brewer(palette = "Spectral", drop=FALSE, na.value='grey80', name='Ponding  Duration') + scale_x_discrete(drop=FALSE) + theme_minimal() + xlab('') + ylab('') + facet_wrap(vars(groupname)) + theme(legend.position="bottom")
```


## RV Summaries by Depth
Median component RV along 1cm slices, grouped by component name and mapunit type ("new" vs. "old").
```{r echo=FALSE, results='hide', fig.width=slab.fig.width, fig.height=6, out.width="100%"}
# quantiles over depth slices as grouped by dmuname/vintage labels
a <- slab(x, dmuname ~ claytotal_r + cec7_r + awc_r + om_r + ph1to1h2o_r, slab.fun = aqp:::.slab.fun.numeric.fast)
a$dmuname <- factor(a$dmuname)

# better plot style
cols <- brewer.pal(n=8, 'Paired')[c(2,1,4,3,6,5,8,7,10,9)]
tps <- list(superpose.line=list(lwd=2, col=cols))

xyplot(
  top ~ p.q50 | variable, data=a, groups=dmuname,
  lower=a$p.q5, upper=a$p.q95, sync.colors=TRUE, alpha=0.5,
  ylim=c(160,-5), layout=c(5,1), scales=list(x=list(relation='free')),
  xlab='', ylab='',
  par.settings=tps,
  strip = strip.custom(bg=grey(0.85)),
  panel=panel.depth_function, 
  prepanel=prepanel.depth_function,
  auto.key=list(columns=length(levels(a$dmuname)), lines=TRUE, points=FALSE)
)
```

Median component RV along 1cm slices, grouped by component name and mapunit type ("new" vs. "old"), and split into panels by national map unit symbol.
```{r echo=FALSE, results='hide', fig.width=slab.fig.width, fig.height=mu.prop.panels.height, out.width="100%"}
# iterate over MU
mu <- unique(x$nationalmusym)
ll <- list()
for(i in mu) {
  
  # subset to current MU
  # quantiles over depth slices as grouped by dmuname/vintage labels
  # just in case there are multiple "old" / "new" groups, use quantiles vs. identity
  a <- slab(x[which(x$nationalmusym == i), ], dmuname ~ claytotal_r + cec7_r + awc_r + om_r + ph1to1h2o_r, slab.fun = aqp:::.slab.fun.numeric.fast)
  
  # convert to factor for plotting
  a$dmuname <- factor(a$dmuname)
  
  # add to indexed list
  ll[[i]] <- a
}

# convert back to DF
aa <- ldply(ll)

# plot styling
cols <- brewer.pal(n=8, 'Paired')[c(2,1,4,3,6,5,8,7,10,9)]
tps <- list(superpose.line=list(lwd=2, col=cols))

p <- xyplot(
  top ~ p.q50 | variable + .id, data=aa, groups=dmuname,
  ylim=c(200,-5), 
  scales=list(x=list(relation='free'), alternating=3),
  par.settings=tps,
  panel=panel.depth_function, 
  prepanel=prepanel.depth_function,
  auto.key=list(columns=length(levels(a$dmuname)), lines=TRUE, points=FALSE)
)

# move MU groups to outerstrip
p <- useOuterStrips(p, strip = strip.custom(bg=grey(0.85)), strip.left = strip.custom(bg=grey(0.75)))

# fix axis labels and print
update(p, xlab='', ylab='')
```


----------------------------
This document is based on `soilDB` version `r utils::packageDescription("soilDB", field="Version")` and `aqp` version `r utils::packageDescription("aqp", field="Version")`.
<br>
Report [configuration and source code are hosted on GitHub](https://github.com/ncss-tech/soilReports).


