---
pagetitle: 'DMU Summary'
output:
  html_document:
    mathjax: null
    smart: no
    keep_md: no
params:
  musym: '3110'
---


```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
	.report.name <- 'DMU-summary'
	.report.version <- '0.2'
	.report.description <- 'DMU Summary Report'
```

```{r echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(knitr, quietly=TRUE)

# chunk options
opts_chunk$set(message=FALSE, warning=FALSE, background='#F7F7F7', fig.align='center', fig.retina=2, dev='png', tidy=FALSE, verbose=FALSE, progress=FALSE, echo = FALSE)


library(aqp, quietly=TRUE)
library(soilDB, quietly=TRUE)
library(sharpshootR, quietly=TRUE)
library(latticeExtra, quietly=TRUE)
library(reshape2, quietly=TRUE)
library(tactile, quietly=TRUE)
library(ggplot2, quietly=TRUE)
library(cluster, quietly=TRUE)

source('custom.R')


## re-make cached data
# source('cache-data.R')


# load cached data
load('data.rda')

## debugging
# params <- list(musym = '3145')

## subset pieces

# component + mu records (SPC)
co <- subset(co, musym == params$musym)

# component month (DF)
cm <- subset(cm, subset = coiid %in% profile_id(co))


# component pedon linkage (DF)
cp <- subset(cp, subset = coiid %in% profile_id(co))

# pedons (SPC)
p <- subset(p, profile_id(p) %in% cp$peiid)

# subset extended OSD data here
osds.ac <- subset(osds$climate.annual, subset = series %in% toupper(unique(co$compname)))

# just the SPC
osds <- subset(osds$SPC, profile_id(osds$SPC) %in% toupper(unique(co$compname)))

cotx <- subset(cotx, subset = coiid %in% profile_id(co))

geom <- subset(geom, subset = coiid %in% profile_id(co))

pm <- subset(pm, subset = coiid %in% profile_id(co))


## re-level component labels
co$.label <- factor(
  co$.label,
  levels = co$.label[order(co$comppct_r, decreasing = TRUE)]
)


## add component pedon / component data to pedons

## TODO: check to make sure that there is only a single case of each pedon
any(table(cp$peiid) > 1)


# merge subset component pedons into SPC
site(p) <- cp[, c('peiid', 'coiid', 'representative')]


# look-up associated component label
site(p)$.comp_label <- co$.label[match(p$coiid, site(co)$coiid)]

## TODO: sort component names / labels by decreasing component percent

# pedon convenience label
p$.pedon_label <- p$taxonname

# flag RV copedon
idx <- which(p$representative == 1)
p$.pedon_label[idx] <- paste0(p$.pedon_label[idx], '|R')

# flag OSD copedon
idx <- which(p$pedontype == 'OSD pedon')
p$.pedon_label[idx] <- paste0(p$.pedon_label[idx], '|O')


## establish some reasonable figure widths in inches

# copedon figure
copedon.profile.fig.width <- 2 + (length(p) * 1.5)

# component RV thematic sketches
comp.profile.fig.width <- 2 + (length(co) * 1.5)

# OSD dendrogram for all components that are named series
# this figure is only possible if there are more than 1 series
do.osd.dend <- TRUE
osd.fig.width <- 5 + ((length(co) + length(osds)) * 0.7)

# component comparison by profile_compare
do.comp.comparison <- length(co) > 1

```




<br>
<div style="text-align: center; border-top-style: solid; border-bottom-style: solid; border-top-width: 2px; border-bottom-width: 2px;">
<span style="font-size: 150%; font-weight: bold;">
`r sprintf("%s: %s", co$musym[1], co$muname[1])`
</span>
<br>
report version `r .report.version`
<br>
`r format(Sys.time(), "%Y-%m-%d %H:%M")`</div>
<hr>


<br>
This report requires setting up a selected set in your local NASIS database:

   * load legend by area symbol
   * load related MU (approved / provisional only)
   * load related DMU (rep DMU only)
   * load related component pedons
   * load related site observation


## Components
```{r echo = FALSE}
co.summary <- site(co)[order(co$comppct_r, decreasing = TRUE), c('musym', '.label', 'compkind', 'taxclname', 'hydgrp', 'ecosite_name')]

kableExtra::kable_styling(
  knitr::kable(co.summary, row.names = FALSE, format = 'html'),
   full_width = TRUE, font_size = 11
)
```


### Component Parent Material | Landform
```{r echo = FALSE}
pm.summary <- site(co)[, c('.label', 'landform_string', 'pmkind', 'pmorigin')]

kableExtra::kable_styling(
  knitr::kable(pm.summary, row.names = FALSE, format = 'html'),
   full_width = FALSE
)
```


### Component Text Notes
```{r echo = FALSE}
txt <- cotx[which(cotx$textcat == 'GENSOIL'), ]

txt <- merge(
  site(co)[, c('coiid', '.label', 'comppct_r')],
  txt,
  by = 'coiid',
  all.x = TRUE,
  sort = FALSE
)

txt <- txt[order(txt$comppct_r, decreasing = TRUE), ]

kableExtra::kable_styling(
  kable(txt[, c('.label', 'textentry')], row.names = FALSE, format = 'html'), full_width = FALSE, font_size = 11
)
```

<!-- ### Pedon Data Usage / Borrowing -->
<!-- too much output -->
```{r eval = FALSE, echo = FALSE, results='asis'}
txt <- cotx[which(cotx$textcat == 'rep pedon'), ]

txt <- merge(
  site(co)[, c('coiid', '.label', 'comppct_r')],
  txt,
  by = 'coiid',
  all.x = TRUE,
  sort = FALSE
)

txt <- txt[order(txt$comppct_r, decreasing = TRUE), ]

# # attempt to read-in as white-space delimited
# # hmm not going to work, without some clever tricks
# nm <- read.table(textConnection(object = txt$textentry[1]), nrows = 1)
# 
# txt.table <- read.table(textConnection(object = txt$textentry[1]), skip = 3)
# names(txt.table) <- nm


for(i in 1:nrow(txt)) {
  cat('<div id="mu_textnote"><pre>')
  cat(txt$.label[i])
  cat('<hr>')
  cat(txt$textentry[i], sep = '\n')
  cat('</prev></div>')
}

```



<!-- TODO: this is failing -->
<!-- climate comparison / summary between OSDs (if present and component) -->
```{r echo=FALSE, results='hide', fig.width=12, fig.height=6.5, eval=do.osd.dend}
try(
  print(vizAnnualClimate(osds.ac)$fig)
  )
```

### Component Climate
```{r echo = FALSE, fig.width=10, fig.height=8}
# comp climate data
nm <- siteNames(co)[grep('^ffd|^maat|^map|^slope|^elev', siteNames(co))]
co.climate <- site(co)[, c('.label', nm)]

# melt
co.climate.long_r <- melt(co.climate, id.vars = '.label', measure.vars = nm[grep('_r', nm)])
co.climate.long_l <- melt(co.climate, id.vars = '.label', measure.vars = nm[grep('_l', nm)])
co.climate.long_h <- melt(co.climate, id.vars = '.label', measure.vars = nm[grep('_h', nm)])

# fix names
names(co.climate.long_r)[3] <- 'rv'

# remove _r
co.climate.long_r$variable <- gsub(pattern = '_r', replacement = '', x = co.climate.long_r$variable, fixed = TRUE)

# row-order is preserved, copy over low / high
co.climate.long_r$low <- co.climate.long_l$value
co.climate.long_r$high <- co.climate.long_h$value

# factor levels / units
co.climate.long_r$variable <- factor(
  co.climate.long_r$variable, 
  levels = c('elev', 'ffd', 'maat', 'map', 'slope'),
  labels = c('Elevation (m)', 'Frost-Free Days', 'MAAT (deg C)', 'MAP (mm)', 'Slope (%)'))


## TODO: generalize this
# add area-wide range
.local_range <- data.frame(
  .label = 'CA792',
  variable = factor(
    c('elev', 'ffd', 'maat', 'map', 'slope'),
    labels = c('Elevation (m)', 'Frost-Free Days', 'MAAT (deg C)', 'MAP (mm)', 'Slope (%)')
    ),
  rv = c(3005, 83, 3.28, 984, 43),
  low = c(1469, 40, -1.02, 583, 8),
  high = c(3728, 193, 12.01, 1381, 99)
  
)

# combine components with area-wide ranges
.combined <- rbind(co.climate.long_r, .local_range)
.combined$.label <- factor(.combined$.label, levels = c(levels(co.climate.long_l$.label), unique(.local_range$.label)))

segplot(
  factor(.label) ~ low + high | variable, 
  data = .combined, 
  centers = rv, 
  draw.bands = FALSE, 
  scales = list(x = list(relation = 'free', rot = 45)),
  strip = strip.custom(bg = grey(0.85)),
  lwd = 2, pch = 15, cex = 1.25,
  par.settings = tactile.theme(),
  layout = c(3, 2),
  as.table = TRUE,
  panel = function(...) {
    panel.grid(h = -1, v = -1)
    panel.segplot(...)
  }
  )
```



### Component | OSD Taxonomic Comparison
```{r echo=FALSE, results='hide', fig.width=osd.fig.width, fig.height=6, eval=do.osd.dend}
# combine OSDs + comp

co.tmp <- co
site(co.tmp)$soilorder <- co.tmp$taxorder
site(co.tmp)$suborder <- co.tmp$taxsuborder
site(co.tmp)$greatgroup <- co.tmp$taxgrtgroup
site(co.tmp)$subgroup <- co.tmp$taxsubgrp

profile_id(co.tmp) <- as.character(co.tmp$.label)
co.tmp <- combine(co.tmp, osds)

# wrapping in try(): there are some cases where there aren't enough data
#                    to compute distance matrix
try(
  SoilTaxonomyDendrogram(co.tmp, cex.taxon.labels = 0.8, width = 0.25, name.style = 'center-center', hz.depths = TRUE, plot.depth.axis = FALSE), 
  silent = TRUE
)

# SoilTaxonomyDendrogram(osds, cex.taxon.labels = 0.8, width = 0.25, name.style = 'center-center', hz.depths = TRUE, plot.depth.axis = FALSE)
```



### Component Diagnostic Features
```{r echo = FALSE, fig.width=8, fig.height=6}
d.wide <- soilDB:::.diagHzLongtoWide(diagnostic_hz(co), feature = 'featkind', id = 'coiid')
co.tmp <- co
site(co.tmp) <- d.wide

## TODO: something wrong when all data share the same features, likely the re-ordering of vars

try(
  diagnosticPropertyPlot2(co.tmp, v = names(d.wide)[-1], k = 3, grid.label = '.label')
)
```


### Component Comparison
```{r echo = FALSE, fig.width=8, fig.height=6, eval = do.comp.comparison}
suppressMessages(
  d <- profile_compare(
    co, 
    vars = c('sandtotal_r', 'claytotal_r', 'fragvoltot_r', 'ph1to1h2o_r', 'om_r'),
    max_d = max(co, v = 'claytotal_r'),
    k = 0,
    rescale.result = TRUE
  )
)

h <- as.hclust(diana(d))

par(mar = c(0, 0, 0, 0))
plotProfileDendrogram(co, clust = h, width = 0.25, name.style = 'center-center', hz.depths = TRUE, plot.depth.axis = FALSE, label = '.label', y.offset = 0.15, scaling.factor = 0.007, color = 'claytotal_r')
```

<!-- texture triangle by component -->
```{r echo = FALSE, eval = FALSE}
textureTriangleSummary(data.frame(SAND=co$sandtotal_r, SILT=co$silttotal_r, CLAY = co$claytotal_r))
```


### Component Pedons
```{r echo = FALSE, fig.width = copedon.profile.fig.width, fig.height = 6.5}
par(mar = c(0, 0, 1, 0))
# component convenience labels
groupedProfilePlot(p, groups = '.comp_label', label = '.pedon_label', group.name.offset = c(-15, -5), id.style = 'side', group.name.cex = 0.75, name.style = 'center-center', hz.depths = TRUE, plot.depth.axis = FALSE, cex.names = 0.66)
```



### Component RV Evaluation
```{r echo = FALSE, fig.width=comp.profile.fig.width, fig.height=6.5}
par(mar=c(0.25, 0.5, 4, 0))

plotSPC(co, label='.label', group.name.cex = 0.75, color='texture', col.label='Texture', col.legend.cex=0.75, name.style = 'center-center', hz.depths = TRUE, plot.depth.axis = FALSE, cex.names = 0.66, width = 0.25)

plotSPC(co, label='.label', group.name.cex = 0.75, color='ksat_r', col.label='Ksat (\U00B5m/sec)', col.legend.cex=0.75, name.style = 'center-center', hz.depths = TRUE, plot.depth.axis = FALSE, cex.names = 0.66, width = 0.25)

plotSPC(co, label='.label', group.name.cex = 0.75, color='om_r', col.label='Organic Matter (%)', col.legend.cex=0.75, name.style = 'center-center', hz.depths = TRUE, plot.depth.axis = FALSE, cex.names = 0.66, width = 0.25)

plotSPC(co, label='.label', group.name.cex = 0.75, color='lep_r', col.label='LEP', col.legend.cex=0.75, name.style = 'center-center', hz.depths = TRUE, plot.depth.axis = FALSE, cex.names = 0.66, width = 0.25)

plotSPC(co, label='.label', group.name.cex = 0.75, color='cec7_r', col.label='CEC @ pH 7 (cmol[+] / kg)', col.legend.cex=0.75, name.style = 'center-center', hz.depths = TRUE, plot.depth.axis = FALSE, cex.names = 0.66, width = 0.25)

plotSPC(co, label='.label', group.name.cex = 0.75, color='ecec_r', col.label='ECEC (cmol[+] / kg)', col.legend.cex=0.75, name.style = 'center-center', hz.depths = TRUE, plot.depth.axis = FALSE, cex.names = 0.66, width = 0.25)

plotSPC(co, label='.label', group.name.cex = 0.75, color='sumbases_r', col.label='Sum of Bases (cmol[+] / kg)', col.legend.cex=0.75, name.style = 'center-center', hz.depths = TRUE, plot.depth.axis = FALSE, cex.names = 0.66, width = 0.25)
```


### Component Low/RV/High Evaluation
```{r echo = FALSE, fig.width = 10, fig.height = 6.5}
# iterate over component IDs
# decreasing component percentage order
co.ids <- site(co)$coiid[order(co$comppct_r, decreasing = TRUE)]
for(comp in co.ids){
  
  ## subset components: component record ID
  co.sub <- subset(co, site(co)$coiid == comp)
  
  ## subset pedons: component record ID
  p.sub <- subset(p, coiid == comp)
  
  ## OSDs: component name
  co.name <- co$compname[match(comp, site(co)$coiid)]
  osds.sub <- subset(osds, id == toupper(co.name))
  
  # there may not be an OSD yet
  if(length(osds.sub) == 0) {
    # use filler based on deepest component subset
    osds.sub <- emtpySPC(co.sub[1, ], top = 0, bottom = max(co.sub)) 

  }
  
  # there may be no component pedons
  if(length(p.sub) == 0) {
    # use filler based on deepest component subset
    p.sub <- emtpySPC(co.sub[1, ], top = 0, bottom = max(co.sub)) 
  }


## overview sketches
OverviewSketches(osds.sub, co.sub, p.sub)

## thematic sketches

# v.co = component basename
# v.p = pedon name

thematicSketches(v.co = 'claytotal', v.p = 'clay', fig.title = 'Clay Content (%)', osds.sub, co.sub, p.sub)

thematicSketches(v.co = 'sandtotal', v.p = 'sand', fig.title = 'Sand Content (%)', osds.sub, co.sub, p.sub)

thematicSketches(v.co = 'fragvoltot', v.p = 'total_frags_pct', fig.title = 'Total Fragments (%)', osds.sub, co.sub, p.sub)

thematicSketches(v.co = 'ph1to1h2o', v.p = 'phfield', fig.title = 'pH 1:1 H2O', osds.sub, co.sub, p.sub)


} # end iterating over components

```





### Component Month
```{r echo=FALSE, results='hide', fig.width=9, fig.height=4.5}
ggplot(cm, aes(month, .label, flodfreqcl)) + geom_tile(aes(fill = flodfreqcl), color='white', lwd=1) + scale_fill_brewer(palette = "Spectral", drop=FALSE, na.value='grey80', name='Flooding Frequency') + scale_x_discrete(drop=FALSE) + theme_minimal() + xlab('') + ylab('') + theme(legend.position="bottom") + guides(fill=guide_legend(nrow=2,byrow=TRUE))

ggplot(cm, aes(month, .label, floddurcl)) + geom_tile(aes(fill = floddurcl), color='white', lwd=1) + scale_fill_brewer(palette = "Spectral", drop=FALSE, na.value='grey80', name='Flooding Duration') + scale_x_discrete(drop=FALSE) + theme_minimal() + xlab('') + ylab('') + theme(legend.position="bottom") + guides(fill=guide_legend(nrow=2,byrow=TRUE))

ggplot(cm, aes(month, .label, pondfreqcl)) + geom_tile(aes(fill = pondfreqcl), color='white', lwd=1) + scale_fill_brewer(palette = "Spectral", drop=FALSE, na.value='grey80', name='Ponding  Frequency') + scale_x_discrete(drop=FALSE) + theme_minimal() + xlab('') + ylab('') + theme(legend.position="bottom") + guides(fill=guide_legend(nrow=2,byrow=TRUE))

ggplot(cm, aes(month, .label, ponddurcl)) + geom_tile(aes(fill = ponddurcl), color='white', lwd=1) + scale_fill_brewer(palette = "Spectral", drop=FALSE, na.value='grey80', name='Ponding  Duration') + scale_x_discrete(drop=FALSE) + theme_minimal() + xlab('') + ylab('') + theme(legend.position="bottom") + guides(fill=guide_legend(nrow=2,byrow=TRUE))
```


