---
title: "Interactive Component Report - static export"
output:
  html_document:
  mathjax: null
jquery: null
smart: no
keep_md: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = F,
  message = FALSE,
  warning = FALSE,
  fig.align = 'center',
  out.width = "100%",
  fig.retina = 2,
  dev = 'png',
  antialias = 'cleartype'
)
# LOAD PACKAGES (used in config.R, shiny.Rmd and report.Rmd)
source("packages.R")
```

Shiny Pedon Summary
==================
<br>
`r format(Sys.time(), "%Y-%m-%d")`
<br>
<center>**`r input$reportname`**

ranges are (`r p.low.rv.high`) percentiles
</center>

### Taxon Names and Pedon Types
Check to make sure that pedons used within this report have been correctly assigned to this component. 
```{r pedon-kind-tables, echo=FALSE, results='asis'}
# join missing genhz data to site
f <- peds
site(f) <- s$mgz
site.i <- site(f)

kable(site.i[order(site.i$musym, site.i$upedonid), c('musym', 'upedonid', 'taxonname', 'taxonkind', 'pedontype', 'bedrckdepth', 'taxpartsize', 'taxsubgrp', 'missing.genhz')], row.names = FALSE, align = 'c')
```

## MUSYM: `r input$s.mu`; Component: `r input$component_name`; Modal: `r input$modal_pedon`

### Grouped Profile Plot
```{r,fig.width=12, fig.height=8}
#TODO: abstract these with "reactive" and "static" available as a toggle
peds$taxonname <- factor(peds$taxonname)
groupedProfilePlot(peds, groups = 'taxonname', label = 'upedonid',
                   print.id = TRUE, id.style = 'side', cex.id = 1.2,
                   cex.names = 1,
                   depth.axis = list(cex = 1.25, line = -3.0),
                   y.offset = 7, group.name.cex = 1,
                   group.name.offset = -8, color = input$thematic_field,
                   width = 0.1, shrink = T, shrink.cutoff = 3)
```

### Map

```{r width=40, height=30}
pedon_locations <- sf::st_as_sf(as(peds,'SpatialPointsDataFrame'))
      slot(mapView(pedon_locations, height=2000, width=2000, label=pedon_locations$upedonid, map.types=c("Stamen.TopOSMRelief",
                                                "Esri.WorldTopoMap",
                                                "Esri.WorldImagery",
                                                "OpenStreetMap.Mapnik")), 'map') 
```

### Slab-wise Profile plot
Blue line shows the median slab value for the selected set of pedons, with the 5th to 95th percentile envelope shown in gray. Thick red line shows the values from selected modal pedon.
```{r,fig.width=8, fig.height=8}
if(!input$thematic_field %in% c("moist_soil_color","dry_soil_color")) {
  sl <- slab(peds, fm = as.formula(paste0(" ~ ",input$thematic_field)))
  a <- xyplot(top ~ p.q50, data=sl, ylab='Depth',
              xlab=paste0(input$thematic_field,'\nmedian bounded by 5th and 95th percentiles'),
              lower=sl$p.q5, upper=sl$p.q95, ylim=c(250,-5),
              panel=panel.depth_function, 
              prepanel=prepanel.depth_function,
              cf=sl$contributing_fraction, aspect = 1.5,
              layout=c(1,1), scales=list(x=list(alternating=1)))

  if(!is.na(input$modal_pedon)) {
    modalped <- horizons(peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$upedonid,sep=":")),]) 
    #TODO: select modal pedon, need to query additional component data? or use dropdown box of available peds
    modalped$middepth <- modalped$hzdept + (modalped$hzdepb - modalped$hzdept)/2
    
    foo <- 1:(length(modalped$middepth)*3)
    modalpedplot <- data.frame(foo)
    modalpedplot$y <- 1:length(foo)
    modalpedplot$y[which((foo %% 3)==1)] <- modalped$hzdept
    modalpedplot$y[which((foo %% 3)==2)] <- modalped$middepth
    modalpedplot$y[which((foo %% 3)==0)] <- modalped$hzdepb
    modalpedplot$x <- rep(modalped[[input$thematic_field]],each=3)
    b <- xyplot(y~x,data=modalpedplot, type="l", col="RED",lwd=2,ylim=c(250,-5),layout=c(1,1), scales=list(x=list(alternating=1)),par.settings = list(superpose.line = list(lwd=3)))
    (a + as.layer(b))
  } else {
    a
  }
}
```

### Generalized Grouped Profile Plot
```{r,fig.width=12, fig.height=8}
peds$taxonname <- factor(peds$taxonname)
groupedProfilePlot(peds, name = 'genhzraw', groups = 'taxonname', label = 'upedonid',
                   print.id = TRUE, id.style = 'side', cex.id = 1.2,
                   cex.names = 1,  depth.axis = list(cex = 1.25, line = -3.0),
                   y.offset = 7, group.name.cex = 1,
                   group.name.offset = -8, color = input$thematic_field,
                   width = 0.1, shrink = T, shrink.cutoff = 3)
```

### Generalized Horizon Probability
```{r,fig.width=10, fig.height=8}
update(s$ml.hz.plot, aspect = 1.5)
```

### Geomorphology

#### Hillslope position (2D)
```{r}
df <- categorical.prop.table(peds$hillslopeprof)
kable(df)

if(!is.na(input$modal_pedon)) {
  modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$upedonid,sep=":")),] 
  print(paste0("Modal pedon value: ",modal$hillslopeprof))
}
```

#### Geomorphic position - Hills (3D)
```{r}
df <- categorical.prop.table(peds$geomposhill)
kable(df)

if(!is.na(input$modal_pedon)) {
  modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$upedonid,sep=":")),] 
  print(paste0("Modal pedon value: ",modal$geomposhill))
}
```

#### Geomorphic position - Mountains (3D)
```{r}
df <- categorical.prop.table(peds$geomposmntn)
kable(df)

if(!is.na(input$modal_pedon)) {
  modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$upedonid,sep=":")),] 
  print(paste0("Modal pedon value: ",modal$geomposmntn))
}
```

#### Geomorphic position - Flats (3D)
```{r}
df <- categorical.prop.table(peds$geomposflats)
kable(df)

if(!is.na(input$modal_pedon)) {
  modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$upedonid,sep=":")),] 
  print(paste0("Modal pedon value: ",modal$geomposflats))
}
```

#### Geomorphons
```{r}
df <- categorical.prop.table(peds$gis_geomorphons)
kable(df)

if(!is.na(input$modal_pedon)) {
  modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$upedonid,sep=":")),] 
  print(paste0("Modal pedon value: ",modal$gis_geomorphons))
}
```

#### Drainage class
```{r}
df <- categorical.prop.table(peds$drainagecl)
kable(df)

if(!is.na(input$modal_pedon)) {
  modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$upedonid,sep=":")),] 
  print(paste0("Modal pedon value: ",modal$drainagecl))
}
```

#### Surface Shape (DOWN/ACROSS)
```{r}
down <- factor(as.character(peds$shapedown),labels =  c("V","L","C"), levels=c("convex","linear","concave"))
acro <- factor(as.character(peds$shapeacross),labels =  c("V","L","C"), levels=c("convex","linear","concave"))
shape <- factor(paste(as.character(down),as.character(acro),sep="/"))
shape[grepl(shape,pattern="NA")] <- NA
shape <- factor(shape)
df <- categorical.prop.table(shape)
kable(df)

if(!is.na(input$modal_pedon)) {
  modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$upedonid,sep=":")),] 
  print(paste0("Modal pedon value: ",modal$shapedown,modal$shapeacross))
}
```

### Aspect
```{r,fig.width=8, fig.height=8}
if(sum(!is.na(peds$aspect)) > 2)
  aspect.plot(peds$aspect, q=p.low.rv.high, plot.title=input$pedon_pattern, pch=21, bg='RoyalBlue', col='black', arrow.col=c('grey', 'red', 'grey'))
if(!is.na(input$modal_pedon)) {
  modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$upedonid,sep=":")),] 
  print(paste0("Modal pedon value: ",modal$aspect))
}
```

### Ecology

#### Ecological site
```{r}
df <-  categorical.prop.table(peds$ecositeid)
kable(df)

if(!is.na(input$modal_pedon)) {
  modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$upedonid,sep=":")),] 
  print(paste0("Modal pedon value: ",modal$ecositeid))
}
```

#### Plant list
TODO: Print component plant data

### Horizon

#### Field-described versus Generalized Horizonation
```{r} 
    if(length(peds)) {
      m <- as.matrix(table(peds$hzname,peds$genhz))
      colz <- colnames(m)
      rowz <- rownames(m)
      x <- length(colz)
      y <- length(rowz)
      buf <- data.frame(field.hz=rowz)
      for(i in 1:x) {
        buf <- cbind(buf, data.frame(m[,i]))
      }
      colnames(buf) <- c("Field HZ", colz)
      rownames(buf) <- NULL
      DT::datatable(buf, options = list(
        pageLength = 250
      ))
    }
#kable(table(peds$genhz,peds$hzname))
```

#### Horizonation Network Graph
```{r,fig.width=6, fig.height=6}
m <- genhzTableToAdjMat(table(peds$genhz,peds$hzname))
plotSoilRelationGraph(m, graph.mode = 'directed', edge.arrow.size=0.5, vertex.label.family='sans')
```

### Texture
```{r}
kable(s$tt)
#texture.triangle.low.rv.high(data.frame(sand=peds$sandtotest,silt=peds$silttotest,clay=peds$claytotest), p=c(0.05, 0.5, 0.95))  
```

### Color
```{r,fig.width=12, fig.height=8}
aggregateColorPlot(aggregateColor(peds, groups = 'genhz', col = 'soil_color'), label.font = 1, label.cex = 0.95, print.n.hz = TRUE)
```

### Morphology
```{r}
kable(s$rt)
```

### Surface Fragments
```{r}
kable(s$sf)
```

### Diagnostics
```{r}
kable(s$dt)
```

### Diagnostics plot
```{r,fig.width=12, fig.height=10}
 if(length(peds) & nrow(diagnostic_hz(peds))) {
      #return(diagnostic_hz(peds))
      v.possible <- c("anthropic epipedon","abrupt textural change",
              "andic soil properties","cambic horizon","calcic horizon","durinodes",
              "petrocalcic horizon",
             "fragipan","gypsic horizon","gilgai","glossic horizon","aquic conditions",
             "histic epipedon","sombric horizon","petrogypsic horizon","placic horizon",
             "kandic horizon","lithic contact","secondary carbonates","mollic epipedon",
             "melanic epipedon","natric horizon","ochric epipedon","plaggen epipedon",
             "petroferric contact","permafrost","plinthite","albic horizon",
             "agric horizon","spodic horizon","slickensides","argillic horizon",
             "umbric epipedon","sulfuric horizon","paralithic contact","oxic horizon",
             "salic horizon","duripan","sulfidic materials",
             "interfingering of albic materials","fibric soil materials",
             "hemic soil materials","sapric soil materials","humilluvic material",
             "limnic materials","coprogenous earth","diatomaceous earth","marl",
             "albic materials","mottles with chroma 2 or less","lamellae",
             "ortstein","fragic soil properties","densic contact","densic materials",
             "paralithic materials","anhydrous conditions","cryoturbation",
             "gelic materials","glacic layer","folistic epipedon",
             "redox concentrations","redox depletions with chroma 2 or less",
             "reduced matrix","endosaturation","episaturation","anthric saturation",
             "lithologic discontinuity","strongly contrasting particle size class",
             "gypsum accumulations","salt accumulations","volcanic glass",
             "n value > 0.7","spodic materials","artifacts","fibers",
             "free carbonates","resistant minerals","weatherable minerals",
             "anhydritic horizon","human-altered material","human-transported material",
             "manufactured layer","manufactured layer contact")
      v.use <- gsub(v.possible, pattern = "\\s{1}", replacement=".")
      v <- v.use[v.use %in% siteNames(peds)]
      diagnosticPropertyPlot(peds, v = v, k = 2)
    } else {
      print("No pedons matching criteria or diagnostic features populated.")
    }
```

### Mapunit Summary
```{r fig.height=30, fig.width=15}
print(s$pmg)
```

-----------------------------------------
  This document is based on `aqp` version `r utils::packageDescription("aqp", field="Version")` and `soilDB` version `r utils::packageDescription("soilDB", field="Version")`. Interface based on `shiny` version `r utils::packageDescription("shiny", field="Version")` and `rmarkdown` version `r utils::packageDescription("rmarkdown", field="Version")`.
