---
title: "Interactive Component Report - static export"
output:
  html_document:
  mathjax: null
jquery: null
smart: no
keep_md: no
---
<!-- shiny-pedon-summary (v1.0) -- instance created 2020-01-27 14:16:49-->  

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
	.report.name <- 'shiny-pedon-summary'
	.report.version <- '1.0'
	.report.description <- 'Interactively subset and summarize NASIS pedon data from one or more map units'
```
  <!-- shiny-pedon-summary (v0.2) -- instance created 2018-01-03 08:11:25-->  
  
```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
.report.name <- 'shiny-pedon-summary'
.report.version <- '1.0'
.report.description <- 'interactively subset and summarize NASIS pedon data from one or more map units'
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message=FALSE, warning = FALSE, fig.align='center', fig.retina=2, dev='png', antialias='cleartype')
```

Shiny Pedon Summary
==================
<br>
`r format(Sys.time(), "%Y-%m-%d")`
<br>
<center>**`r input$reportname`**

ranges are (`r p.low.rv.high`) percentiles
</center>

### Taxon Names and Pedon Types
Check to make sure that pedons used within this report have been correctly assigned to this component. 
<p>
```{r pedon-kind-tables, echo=FALSE, results='asis'}
# join missing genhz data to site
f <- peds
site(f) <- s$mgz
site.i <- site(f)

print(xtable(site.i[order(site.i$musym, site.i$pedon_id), c('musym', 'pedon_id', 'taxonname', 'taxonkind', 'pedontype', 'bedrckdepth', 'taxpartsize', 'taxsubgrp', 'missing.genhz')]), type='html', include.rownames=FALSE, table.placement="H", caption.placement="top", html.table.attributes='cellpadding="3" cellspacing="3"')
```
</p>

## MUSYM: `r input$s.mu`; Component: `r input$component_name`; Modal: `r input$modal_pedon`

### Grouped Profile Plot
```{r,fig.width=12, fig.height=8}
#TODO: abstract these with "reactive" and "static" available as a toggle
peds$taxonname <- factor(peds$taxonname)
groupedProfilePlot(peds, groups = 'taxonname', label='pedon_id', 
                   print.id = TRUE, id.style = 'side', cex.id=1.2, 
                   cex.names=1, cex.depth.axis=1.25,y.offset=7, 
                   axis.line.offset=-3.0, group.name.cex=1, 
                   group.name.offset = c(-6,-10), color=input$thematic_field,
                   width=0.1,shrink=T,shrink.cutoff=3)
```

### Slab-wise Profile plot
Blue line shows the median slab value for the selected set of pedons, with the 25th to 75th percentile envelope shown in gray. Thick red line shows the values from selected modal pedon.
```{r,fig.width=8, fig.height=8}
sl <- slab(peds, fm = as.formula(paste0(" ~ ",input$thematic_field)))
a <- xyplot(top ~ p.q50, data=sl, ylab='Depth',
            xlab=paste0(input$thematic_field,'\nmedian bounded by 25th and 75th percentiles'),
            lower=sl$p.q25, upper=sl$p.q75, ylim=c(250,-5),
            panel=panel.depth_function, 
            prepanel=prepanel.depth_function,
            cf=sl$contributing_fraction, aspect = 1.5,
            layout=c(1,1), scales=list(x=list(alternating=1)))

if(!is.na(input$modal_pedon)) {
  modalped <- horizons(peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$pedon_id,sep=":")),]) 
  #TODO: select modal pedon, need to query additional component data? or use dropdown box of available peds
  modalped$middepth <- modalped$hzdept + (modalped$hzdepb - modalped$hzdept)/2
  
  foo <- 1:(length(modalped$middepth)*3)
  modalpedplot <- data.frame(foo)
  modalpedplot$y <- 1:length(foo)
  modalpedplot$y[which((foo %% 3)==1)] <- modalped$hzdept
  modalpedplot$y[which((foo %% 3)==2)] <- modalped$middepth
  modalpedplot$y[which((foo %% 3)==0)] <- modalped$hzdepb
  modalpedplot$x <- rep(modalped[[input$thematic_field]],each=3)
  b <- xyplot(y~x,data=modalpedplot, type="l", col="RED",lwd=2,ylim=c(250,-5),layout=c(1,1), scales=list(x=list(alternating=1)),par.settings = list(superpose.line = list(lwd=3)))
  (a + as.layer(b))
} else {
  a
}
```

### Generalized Grouped Profile Plot
```{r,fig.width=12, fig.height=8}
peds$taxonname <- factor(peds$taxonname)
groupedProfilePlot(peds,name='genhzraw',groups = 'taxonname', label='pedon_id', 
                   print.id = TRUE, id.style = 'side', cex.id=1.2, 
                   cex.names=1, cex.depth.axis=1.25,y.offset=7, 
                   axis.line.offset=-3.0, group.name.cex=1, 
                   group.name.offset = c(-6,-10), color=input$thematic_field,
                   width=0.1,shrink=T,shrink.cutoff=3)
```

### Generalized Horizon Probability
```{r,fig.width=10, fig.height=8}
update(s$ml.hz.plot, aspect = 1.5)
```

### Geomorphology

#### Hillslope position (2D)
```{r}
df <- categorical.prop.table(peds$hillslopeprof)
kable(df)

if(!is.na(input$modal_pedon)) {
  modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$pedon_id,sep=":")),] 
  print(paste0("Modal pedon value: ",modal$hillslopeprof))
}
```

#### Geomorphic position - Hills (3D)
```{r}
df <- categorical.prop.table(peds$geomposhill)
kable(df)

if(!is.na(input$modal_pedon)) {
  modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$pedon_id,sep=":")),] 
  print(paste0("Modal pedon value: ",modal$geomposhill))
}
```

#### Geomorphic position - Mountains (3D)
```{r}
df <- categorical.prop.table(peds$geomposmntn)
kable(df)

if(!is.na(input$modal_pedon)) {
  modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$pedon_id,sep=":")),] 
  print(paste0("Modal pedon value: ",modal$geomposmntn))
}
```

#### Geomorphons
```{r}
df <- categorical.prop.table(peds$gis_geomorphons)
kable(df)

if(!is.na(input$modal_pedon)) {
  modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$pedon_id,sep=":")),] 
  print(paste0("Modal pedon value: ",modal$gis_geomorphons))
}
```

#### Drainage class
```{r}
df <- categorical.prop.table(peds$drainagecl)
kable(df)

if(!is.na(input$modal_pedon)) {
  modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$pedon_id,sep=":")),] 
  print(paste0("Modal pedon value: ",modal$drainagecl))
}
```

#### Surface Shape (DOWN/ACROSS)
```{r}
down <- factor(as.character(peds$shapedown),labels =  c("V","L","C"), levels=c("convex","linear","concave"))
acro <- factor(as.character(peds$shapeacross),labels =  c("V","L","C"), levels=c("convex","linear","concave"))
shape <- factor(paste(as.character(down),as.character(acro),sep="/"))
shape[grepl(shape,pattern="NA")] <- NA
shape <- factor(shape)
df <- categorical.prop.table(shape)
kable(df)

if(!is.na(input$modal_pedon)) {
  modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$pedon_id,sep=":")),] 
  print(paste0("Modal pedon value: ",modal$shapedown,modal$shapeacross))
}
```

### Aspect
```{r,fig.width=8, fig.height=8}
if(nrow(site(peds)) > 2)
  aspect.plot(peds$aspect_field, q=p.low.rv.high, plot.title=input$pedon_pattern, pch=21, bg='RoyalBlue', col='black', arrow.col=c('grey', 'red', 'grey'))
if(!is.na(input$modal_pedon)) {
  modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$pedon_id,sep=":")),] 
  print(paste0("Modal pedon value: ",modal$aspect_field))
}
```

### Ecology

#### Ecological site
```{r}
df <-  categorical.prop.table(peds$ecositeid)
kable(df)

if(!is.na(input$modal_pedon)) {
  modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$pedon_id,sep=":")),] 
  print(paste0("Modal pedon value: ",modal$ecositeid))
}
```

#### Plant list
TODO: Print component plant data

### Horizon

#### Field-described versus Generalized Horizonation
```{r} 
kable(table(peds$genhz,peds$hzname))
```

#### Horizonation Network Graph
```{r,fig.width=6, fig.height=6}
m <- genhzTableToAdjMat(table(peds$genhz,peds$hzname))
plotSoilRelationGraph(m, graph.mode = 'directed', edge.arrow.size=0.5, vertex.label.family='sans')
```

### Texture
```{r}
kable(s$tt)
#texture.triangle.low.rv.high(data.frame(sand=peds$sand,silt=peds$silt,clay=peds$clay), p=c(0.05, 0.5, 0.95))  
```

### Color
```{r,fig.width=12, fig.height=8}
aggregateColorPlot(aggregateColor(peds, groups = 'genhz', col = 'soil_color'), label.font = 1, label.cex = 0.95, print.n.hz = TRUE)
```

### Morphology
```{r}
kable(s$rt)
```

### Surface Fragments
```{r}
kable(s$sf)
```

### Diagnostics
```{r}
kable(s$dt)
```

### Diagnostics plot
```{r,fig.width=12, fig.height=8}
#TODO: why does this break with few pedons? are all relevant diagnostics in here? (NB redox)
# kk = 4
# if(length(peds) != 0) {
#   if(nrow(peds) < 4)
#     kk <- nrow(peds) - 2
#   if(nrow(peds) > 2)
#     diagnosticPropertyPlot2(peds, v=c('lithic.contact', 'paralithic.contact', 'argillic.horizon', 'cambic.horizon', 'ochric.epipedon', 'mollic.epipedon', 'very.shallow', 'shallow', 'mod.deep', 'deep', 'very.deep'), k=kk)
# }
```

### Mapunit Summary
```{r}
print(s$pmg,fig.width=12, fig.height=8)
```

-----------------------------------------
  This document is based on `aqp` version `r utils::packageDescription("aqp", field="Version")` and `soilDB` version `r utils::packageDescription("soilDB", field="Version")`. Interface based on `shiny` version `r utils::packageDescription("shiny", field="Version")` and `rmarkdown` version `r utils::packageDescription("rmarkdown", field="Version")`.
