---
title: "Interactive Component Report"
output: flexdashboard::flex_dashboard
runtime: shiny
---

```{r, echo=FALSE, message=FALSE, results='hide'}
#re-vamped component report 

#projected workflow (not complete)
#      start with NASIS selected set containing all potential pedons (could be entire survey area) and DMU(s) to be populated
#      set MU suite (defines an extended spatial extent with similar geology/climate/geography etc. for borrowing pedons in order to develop ranges
#      set criteria for borrowing (set of simple options/rules, can be defined on a regional basis; defined in terms of sets of attributes and related rules)
#      set target MU (by MUSYM/DMUID) - this will be what the report is generated for
#      get the components being used in the DMU; needs at least a stub record with name for each component (no HZ data required)
#      auto-fill pedon sets for each component using taxon name (could also use other attributes such as depth/drainage) using data from within target MU
#        generate grouped profile plot, allow for user interaction to drop "outliers" and reassign groupings; outliers will be excluded from range calculations
#      generate an "extended set" of pedons using taxon name and indices of similarity from the MU suite extent
#        generate grouped profile plot, allow for user interaction to include similar pedons from MU suite; these will be included in range calculation
#      generate grouped profile plot, clearly showing selected data from within and outside MU
#        calculate provisional ranges and estimate "representativeness" of each pedon WITHIN THE MU extent
#        allow user to select representative pedons (at least one per component in DMU) - need not agree with numerical "representative" index.
#      show representative pedons for each component
#        interactive horizon aggregation (stepwise construction of a base regex pattern, applied to the rep pedon)
#        as pattern for aggregation changes, update a plot of just that component's selected data; highlight horizons that are currently not being aggregated
#        allow pattern to be edited explicitly; or "dumbly" by simply piping hz designations on following user selection
#      once a generalized horizon pattern has been developed allow the patterns to be saved (timestamped) to an output folder
#      generate horizonagg.txt for NASIS import

source('config.R')            #REPORT SETTINGS (many of these can be ported over to Shiny inputs)
source('utility_functions.R') #ABSTRACTED/MODULAR APPLICATION LOGIC
source('main.R')              #LOADS DATA, DOES OVERLAYS, ETC. 
```

Inputs {.sidebar}
-------------------------------------
### INPUT
```{r, echo=FALSE}  
inputPanel(
    #selectInput("s.mu", label = "Select target map unit: ",choices = c("",levels(factor(site(pedons)$MUSYM))), selectize=T)
    textInput("s.mu",label="Enter pattern for matching MUSYM",value=".*") 
)

renderUI( {
  #s.comp <<- getMapunitComponents(input$s.mu)
  s.pedons <<- getMapunitPedons(input$s.mu)
  inputPanel( {textInput("component_name",label="Component name:",value="") })
})

renderUI( {
  inputPanel( {
    textInput("pedon_pattern",label="Pattern for matching pedon taxonname:",value=".*") 
  })
})

renderUI( {
  inputPanel( {
    textInput("phase_pattern",label="Pattern for matching pedon local phase:",value="*") 
  })
})

renderUI( {
  inputPanel({
    foo <- input$s.mu
    selectInput("taxon_kind", label = "Select taxon kind: ",
                choices = c("any","family","series","taxadjunct","taxon above family")) 
    })
})

renderUI( {  
  inputPanel( {
    textInput("upid_pattern",label="Pattern for matching userpedonid:",value="*") 
  })
})

renderUI( {  
  inputPanel( {
    textInput("pedon_list",label="Comma-delimited list of pedons (in lieu of above):",value="") 
  })
})

renderUI( {
  inputPanel({
    foo <- input$s.mu
    peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
    selectInput("modal_pedon", label = "Select modal pedon (peiid:userpedonid): ",
                choices = paste(site(peds)$peiid,site(peds)$pedon_id,sep=":")) 
                #TODO: Is it possible to set default to the pedon currently tagged to component? 
                #rep pedon is not returned by component fetch
    })
})

inputPanel(
  selectInput("thematic_field", label = "Select horizon attribute data to plot: ",
              choices = c("clay","sand","phfield","total_frags_pct","moist_soil_color","dry_soil_color"), selected = "clay")
)

#inputPanel( {
actionButton("refresh_tabular", "Refresh Tabular (TODO)")
actionButton("export_report", "Export Report")
#})

observeEvent(input$refresh_tabular, {
  #TODO: abstract different parts of the data loading in main.R for re-use, add checking to verify new records/update spatial overlays etc
  source('main.R')      
})

observeEvent(input$export_report, {
  s.pedons <<- getMapunitPedons(input$s.mu)
  render(input='report.Rmd', output_file=paste0("output/report_",paste(levels(factor(s.pedons$MUSYM)),collapse="-"),"_",input$component_name,".html"))
})
```

Column {.tabset}
-------------------------------------
### Grouped profile plot
```{r}
renderPlot(
  {
    sourcemu <- input$s.mu #ensures that plotting relies on selected MU (ensures reactive update when mu is changed)
    comp <- s.comp
    peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
    
    groupedProfilePlot(peds,groups = 'taxonname', label='pedon_id', 
                         print.id = TRUE, id.style = 'side', cex.id=1.2, 
                         cex.names=1, cex.depth.axis=1.25,y.offset=7, 
                         axis.line.offset=-3.0, group.name.cex=1, 
                         group.name.offset = c(-6,-10), color=input$thematic_field,
                         width=0.1,shrink=T,shrink.cutoff=3)
      options=list(
        width="100%", height=700
      )
  }
)
```

### Map view
```{r}
renderLeaflet({
    sourcemu <- input$s.mu #ensures that plotting relies on selected MU (ensures reactive update when mu is changed)
    comp <- s.comp
    peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
    pedon_locations <- as(peds,'SpatialPointsDataFrame')
    mapview(pedon_locations)@map #  here we just access the leaflet map slot directly for shiny rendering and use renderLeaflets
})
```

### Slab-wise Profile plot

Blue line shows the median slab value for the selected set of pedons, with the 25th to 75th percentile envelope shown in gray. Thick red line shows the values from selected modal pedon.
```{r}
renderPlot(
  {
    sourcemu <- input$s.mu 
    comp <- s.comp
    peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
    
    s <- slab(peds, fm = as.formula(paste0(" ~ ",input$thematic_field)))
    
    a <- xyplot(top ~ p.q50, data=s, ylab='Depth',
             xlab=paste0(input$thematic_field,'\nmedian bounded by 25th and 75th percentiles'),
             lower=s$p.q25, upper=s$p.q75, ylim=c(250,-5),
             panel=panel.depth_function, 
             prepanel=prepanel.depth_function,
             cf=s$contributing_fraction,
             layout=c(1,1), scales=list(x=list(alternating=1)))
    #b <- slab(peds[1,], fm = as.formula(paste0(" ~ ",input$thematic_field)))
    if(!is.na(input$modal_pedon)) {
      modalped <- horizons(peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$pedon_id,sep=":")),]) 
      
      #TODO: select modal pedon, need to query additional component data? or use dropdown box of available peds
      modalped$middepth <- modalped$hzdept + (modalped$hzdepb - modalped$hzdept)/2
      
      foo <- 1:(length(modalped$middepth)*3)
      modalpedplot <- data.frame(foo)
      modalpedplot$y <- 1:length(foo)
      modalpedplot$y[which((foo %% 3)==1)] <- modalped$hzdept
      modalpedplot$y[which((foo %% 3)==2)] <- modalped$middepth
      modalpedplot$y[which((foo %% 3)==0)] <- modalped$hzdepb
      modalpedplot$x <- rep(modalped[[input$thematic_field]],each=3)
      b <- xyplot(y~x,data=modalpedplot, type="l", col="RED",lwd=2,ylim=c(250,-5),layout=c(1,1), scales=list(x=list(alternating=1)),par.settings = list(superpose.line = list(lwd=3)))
      (a + as.layer(b))
    } else {
      a
    }
  }
)
```

### Generalized GPP
```{r}
renderPlot(
  {
    sourcemu <- input$s.mu 
    comp <- s.comp
    peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
    peds$genhzraw <- as.character(peds$genhz) #this plot function gets mad if name is a factor...
    groupedProfilePlot(peds,name='genhzraw',groups = 'taxonname', label='pedon_id', 
                         print.id = TRUE, id.style = 'side', cex.id=1.2, 
                         cex.names=1, cex.depth.axis=1.25,y.offset=7, 
                         axis.line.offset=-3.0, group.name.cex=1, 
                         group.name.offset = c(-6,-10), color=input$thematic_field,
                         width=0.1,shrink=T,shrink.cutoff=3)
      options=list(
        width="100%", height=700
      )
  }
)
```

### Generalized Hz Probability
```{r}
renderPlot(
  {
    sourcemu <- input$s.mu 
    comp <- s.comp
    peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
    s <- summarize.component(peds)
    s$ml.hz.plot
  }
)
```

### Geomorphology

#### Hillslope position (2D)
```{r}
renderTable( {
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
  df <- categorical.prop.table(peds$hillslopeprof)
  df
}, striped=T)

renderUI( {
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
  if(!is.na(input$modal_pedon)) {
      modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$pedon_id,sep=":")),] 
      print(paste0("Modal pedon value: ",modal$hillslopeprof))
  }
})
```

#### Geomorphic position - Hills (3D)
```{r}
renderTable( {
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
  df <- categorical.prop.table(peds$geomposhill)
  df
}, striped=T)
renderUI( {
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
  if(!is.na(input$modal_pedon)) {
      modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$pedon_id,sep=":")),] 
      print(paste0("Modal pedon value: ",modal$geomposhill))
  }
})
```

#### Geomorphic position - Mountains (3D)
```{r}
renderTable( {
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
  df <- categorical.prop.table(peds$geomposmntn)
  df
}, striped=T)
renderUI( {
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
  if(!is.na(input$modal_pedon)) {
      modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$pedon_id,sep=":")),] 
      print(paste0("Modal pedon value: ",modal$geomposmntn))
  }
})
```

#### Geomorphons
```{r}
renderTable( {
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
  df <- categorical.prop.table(peds$gis_geomorphons)
  df
}, striped=T)
renderUI( {
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
  if(!is.na(input$modal_pedon)) {
      modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$pedon_id,sep=":")),] 
      print(paste0("Modal pedon value: ",modal$gis_geomorphons))
  }
})
```

#### Drainage class
```{r}
renderTable( {
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
  df <- categorical.prop.table(peds$drainagecl)
  df
}, striped=T)

renderUI( {
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
  if(!is.na(input$modal_pedon)) {
      modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$pedon_id,sep=":")),] 
      print(paste0("Modal pedon value: ",modal$drainagecl))
  }
})
```

#### Surface Shape (DOWN/ACROSS)
```{r}
renderTable( {
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
  down <- factor(as.character(peds$shapedown),labels =  c("V","L","C"), levels=c("convex","linear","concave"))
  acro <- factor(as.character(peds$shapeacross),labels =  c("V","L","C"), levels=c("convex","linear","concave"))
  shape <- factor(paste(as.character(down),as.character(acro),sep="/"))
  shape[grepl(shape,pattern="NA")] <- NA
  shape <- factor(shape)
  df <- categorical.prop.table(shape)
  df
}, striped=T)

renderUI( {
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
  if(!is.na(input$modal_pedon)) {
      modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$pedon_id,sep=":")),] 
      print(paste0("Modal pedon value: ",modal$shapedown,modal$shapeacross))
  }
})
```

### Aspect
```{r}
renderPlot({
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
  if(nrow(site(peds)) > 2)
    aspect.plot(peds$aspect_field, q=p.low.rv.high, plot.title=input$pedon_pattern, pch=21, bg='RoyalBlue', col='black', arrow.col=c('grey', 'red', 'grey'))
})

renderUI( {
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
  if(!is.na(input$modal_pedon)) {
      modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$pedon_id,sep=":")),] 
      print(paste0("Modal pedon value: ",modal$aspect_field))
  }
})
```

### Ecology

#### Ecological site
```{r}
renderTable( {
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
  df <-  categorical.prop.table(peds$ecositeid)
  df
}, striped=T)

renderUI( {
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
  if(!is.na(input$modal_pedon)) {
      modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$pedon_id,sep=":")),] 
      print(paste0("Modal pedon value: ",modal$ecositeid))
  }
})
```

#### Plant list
TODO: Print component plant data

### Horizon

#### Field-described versus Generalized Horizonation
```{r} 
renderTable({
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
  xtable(table(peds$genhz,peds$hzname))
})
```

#### Horizonation Network Graph
```{r}
renderPlot({
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern) 
  # convert contingency table -> adj. matrix
  m <- genhzTableToAdjMat(table(peds$genhz,peds$hzname))
  
  plotSoilRelationGraph(m, graph.mode = 'directed', edge.arrow.size=0.5, vertex.label.family='sans')
})
```

### Modal
```{r}

#### Modal pedon (field horizonation : generalized horizonation)
renderPlot({
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
  par(mar=c(1,1,1,1))
  if(!is.na(input$modal_pedon)) {
      modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,site(peds)$pedon_id,sep=":")),] 
      modal$hzagg <- paste0(modal$hzname,":",modal$genhz)
      groupedProfilePlot(modal,name='hzagg',groups="taxonname", label='pedon_id', 
                         print.id = TRUE, id.style = 'side', cex.id=1.2, 
                         cex.names=1, cex.depth.axis=1.25,y.offset=7, 
                         axis.line.offset=-3.0, group.name.cex=1, 
                         group.name.offset = -6, color=input$thematic_field,
                         width=0.1,shrink=T,shrink.cutoff=3)
      options=list(
        width="100%", height=700
      )
  }
})
```

### Texture
```{r}
renderTable({
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
  s <- summarize.component(peds)
  df <- s$tt
  df
})

renderPlot({
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
  
  plot(texture.triangle.low.rv.high(data.frame(sand=peds$sand,silt=peds$silt,clay=peds$clay), p=c(0.05, 0.5, 0.95)))
})
```

### Color
```{r}
renderPlot({
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
  aggregateColorPlot(aggregateColor(peds, groups = 'genhz', col = 'soil_color'), label.font = 2, label.cex = 0.95, print.n.hz = TRUE)
})
```

### Morphology
```{r}
renderTable({
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
  s <- summarize.component(peds)
  df <- s$rt
  df
})
```

### Surface Fragments
```{r}
renderTable({
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
  s <- summarize.component(peds)
  df <- s$sf
  df
})
```

### Diagnostics
```{r}
renderTable({
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
  s <- summarize.component(peds)
  df <- s$dt
  df
})
```

### Diagnostics plot
```{r}
renderPlot({
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
  
  diagnosticPropertyPlot2(peds, v=c('lithic.contact', 'paralithic.contact', 'argillic.horizon', 'cambic.horizon', 'ochric.epipedon', 'mollic.epipedon', 'very.shallow', 'shallow', 'mod.deep', 'deep', 'very.deep'), k=4)
})
```
  
### Mapunit Summary
```{r}
renderPlot({
  sourcemu <- input$s.mu 
  peds <- getPedonsByPattern(input$pedon_pattern,input$upid_pattern,input$pedon_list,input$taxon_kind,input$phase_pattern)
  s <- summarize.component(peds)
  print(s$pmg)
})
```