---
title: "SHINY PEDON SUMMARY"
output: 
  flexdashboard::flex_dashboard
runtime: shiny
---

Inputs {.sidebar data-width=300}
-------------------------------------
### INPUT
```{r, warning=FALSE, echo=FALSE, results='hide', message=FALSE}  
    
# ABSTRACTED APPLICATION LOGIC
source('util.R')

# REPORT SETTINGS 
source('config.R')
```        

```{r, warning=FALSE, echo=FALSE, message=FALSE, results='hide'}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE, fig.align='center', fig.retina=2, dev='png', antialias='cleartype')
                      
if(!exists("loaded"))
  loaded <<- FALSE

res <- loadReportData()

# global variables pedons and pedons_spdf shared between chunks
pedons <<- res$pedons
pedons_spdf <<- res$pedons_spdf
```

```{r, warning=FALSE, echo=FALSE, error=FALSE}
    inputPanel({
        textInput("s.mu",
                  label=sprintf("Pattern [%s]:", musym.col),
                  value=".*")
    })
    inputPanel({
        textInput("report_name",
                  label="Report name:",
                  value=paste0("ReportName_", Sys.Date()))
    })
    inputPanel({
        textInput("pedon_pattern", 
                  label="Pattern [taxonname]:",
                  value=".*")
    })
    inputPanel( {
        textInput("phase_pattern", 
                  label="Pattern [localphase]:",
                  value=".*") 
    })
    inputPanel({
        selectInput("taxon_kind", 
                    label = "Select [taxonkind]: ",
                    choices = c(".*","family","series",
                                "taxadjunct",
                                "taxon above family")) 
    })
  
    inputPanel({
        textInput("upid_pattern",
              label="Pattern [upedonid]:",
              value=".*") 
     })
  
  # renderUI( {  
  #   inputPanel( {
  #     textInput("pedon_list",
  #               label="Comma-delimited list of pedons (in lieu of above):",
  #               value="") 
  #   })
  # })
  if(loaded & !is.null(isolate(input))) {
    #print(isolate(input))
    renderUI(inputPanel({
      peds <- getPedonsByPattern(input, pedons, 
                                  input$s.mu,
                                  input$pedon_pattern, 
                                  input$upid_pattern,
                                  input$pedon_list,
                                  input$taxon_kind,
                                  input$phase_pattern)
       
      selectInput("modal_pedon", 
                   label = "Select modal pedon (peiid:upedonid): ",
                   choices = paste(site(peds)$peiid,
                                   site(peds)$pedon_id,
                                   sep=":")) 
  # TODO: Is it possible to copedon rv
       }))
  }
  
  inputPanel(selectInput("thematic_field", 
              label = "Select horizon data to plot: ",  
              choices = c("clay", "sand", "phfield",
                          "total_frags_pct", 
                          "moist_soil_color",
                           "dry_soil_color"), selected = "clay"))
  
  #inputPanel( {
  actionButton("refresh_tabular", "Refresh Tabular")
  actionButton("export_report", "Export Report")
  #})
  
  observeEvent(input$refresh_tabular, {
    res <- loadReportData()
    
    # global variables pedons and pedons_spdf shared between chunks
    pedons <<- res$pedons
    pedons_spdf <<- res$pedons_spdf
  })
    
  observeEvent(input$export_report, {
    # create output folder if needed
    if(!dir.exists("output"))
      dir.create("output")
    
    # build report environment with user-selected filter settings
    my.env <- new.env()
    sys.source('config.R', my.env)
    sys.source('util.R', my.env)
    
    res <- loadReportData()
    
    my.env$input <- isolate(input)
    my.env$inputreportname <- input$reportname
    my.env$peds <- getPedonsByPattern(input, res$pedons,
                                   input$s.mu,
                                   input$pedon_pattern,
                                   input$upid_pattern,
                                   input$pedon_list,
                                   input$taxon_kind,
                                   input$phase_pattern)
    my.env$s <- summarize.component(my.env$peds)
    
    my.env$peds$genhz <- factor(my.env$peds$genhz, 
                           levels = guessGenHzLevels(res$pedons)$levels)
    
    # render in constructed environment
    render(input='report.Rmd', envir = my.env,
           output_file=paste0("output/report_",
                              input$report_name,".html"))
  })
```

Column {.tabset}
-------------------------------------
### Grouped profile plot
```{r, warning=FALSE, echo=FALSE, error=FALSE}
if(loaded)
  renderPlot({
      if(!is.null(isolate(input)) & length(pedons)) {
        #print(isolate(input))
        peds <- getPedonsByPattern(input, pedons,
                                   input$s.mu,
                                   input$pedon_pattern,
                                   input$upid_pattern,
                                   input$pedon_list,
                                   input$taxon_kind,
                                   input$phase_pattern)
        if(length(peds)) {
          peds$taxonname <- factor(peds$taxonname)
    
          # fix for single-group sets 
          gno <-  c(-6,-10)
          if(length(levels(peds$taxonname)) == 1)
            gno <- -8
            
          groupedProfilePlot(peds, groups = 'taxonname',
                              label='pedon_id',
                              print.id = TRUE, id.style = 'side', cex.id=1.2,
                              cex.names=1, cex.depth.axis=1.25,y.offset=7,
                              axis.line.offset=-3.0, group.name.cex=1,
                              group.name.offset = gno,
                              color=input$thematic_field,
                              width=0.1,shrink=T,shrink.cutoff=3)
            options=list(
              width="100%", height=700
            )  
        }
      }
    })
```

### Map view
```{r, warning=FALSE, echo=FALSE, error=FALSE}
if(loaded)
  renderLeaflet({
  
    peds <- getPedonsByPattern(input, pedons, 
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
    if(length(peds) & 
       inherits(slot(peds,'sp'), 'Spatial') &
        ncol(slot(slot(peds, 'sp'),'coords')) > 1) {
      pedon_locations <- as(peds,'SpatialPointsDataFrame')
      
      #  access leaflet map slot directly for renderLeaflet()
      slot(mapView(pedon_locations, label=pedon_locations$pedon_id, map.types=c("Stamen.TopOSMRelief",
                                                "Esri.WorldTopoMap",
                                                "Esri.WorldImagery",
                                                "OpenStreetMap.Mapnik")), 'map') 
    }
  })
```

### Slab-wise Profile plot

Blue line shows the median slab value for the selected set of pedons, with the 5th to 95th percentile envelope shown in gray. Thick red line shows the values from selected modal pedon. Slabs with less than 1% of pedon data contributing have been omitted for clarity.

```{r, warning=FALSE, echo=FALSE, error=FALSE}
if(loaded)
  renderPlot(
    {
      peds <- getPedonsByPattern(input, pedons, 
                                 input$s.mu,
                                 input$pedon_pattern, 
                                 input$upid_pattern,
                                 input$pedon_list,
                                 input$taxon_kind,
                                 input$phase_pattern)
      #print(profile_id(peds))
      
      if(length(peds) & length(input$thematic_field)) {
        if(!input$thematic_field %in% c("moist_soil_color","dry_soil_color")) {
          
          s <- slab(peds, fm = as.formula(sprintf( " ~ %s", 
                                                  input$thematic_field)))
          # max depth is <1% contributing fraction
          max.idx <- which(s$contributing_fraction <= 0.01)
          if(!length(max.idx))
            max.idx <- nrow(s)
          s.sub <- s[1:max.idx[1],]
          s.sub <- s.sub[complete.cases(s.sub),]
          #print((s.sub))
          
          a <- xyplot(top ~ p.q50, data=s.sub, ylab='Depth',
                      xlab=paste0(input$thematic_field,
                           '\nmedian bounded by 5th and 95th percentiles'), 
                      lower=s$p.q5, upper=s$p.q95, 
                      ylim=c(max(s.sub$bottom, na.rm=TRUE),-5),
                      panel=panel.depth_function, 
                      prepanel=prepanel.depth_function,
                      cf=s$contributing_fraction,
                      layout=c(1,1), scales=list(x=list(alternating=1)))
          
          b <- slab(peds[1,], fm = as.formula(paste0(" ~ ",input$thematic_field)))
  
          if(length(input$modal_pedon) & !is.na(input$modal_pedon)) {
            modalped <- horizons(peds[which(input$modal_pedon == paste(site(peds)$peiid,
                                                                       site(peds)$pedon_id,
                                                                       sep=":")),])
  
            #TODO: select modal pedon from RV copedon
            modalped$middepth <- modalped$hzdept + (modalped$hzdepb - modalped$hzdept)/2
  
            foo <- 1:(length(modalped$middepth)*3)
            modalpedplot <- data.frame(foo)
  
            modalpedplot$y <- 1:length(foo)
            modalpedplot$y[which((foo %% 3) == 1)] <- modalped$hzdept
            modalpedplot$y[which((foo %% 3) == 2)] <- modalped$middepth
            modalpedplot$y[which((foo %% 3) == 0)] <- modalped$hzdepb
  
            modalpedplot$x <- rep(modalped[[input$thematic_field]],each=3)
            b <- xyplot(y ~ x, data=modalpedplot, type="l",
                        col="RED", lwd=3, ylim=c(250,-5), layout=c(1,1),
                        scales=list(x=list(alternating=1)),
                        par.settings = list(superpose.line = list(lwd=3)))
            (a + as.layer(b))
          } else {
            a
          }
        } else {
          print("Color quantiles by depth coming soon")
        }
      }
    }
  )
```

### Generalized GPP
```{r, warning=FALSE, echo=FALSE, error=FALSE}
if(loaded)
  renderPlot(
    {
      peds <- getPedonsByPattern(input, pedons, 
                                 input$s.mu,
                                 input$pedon_pattern, 
                                 input$upid_pattern,
                                 input$pedon_list,
                                 input$taxon_kind,
                                 input$phase_pattern)
      peds$genhz <- factor(peds$genhz, levels = guessGenHzLevels(peds)$levels)
      if(length(peds)) {
        # this plot function gets mad if hzname is a factor...
        # but taxonname must be a factor
        peds$genhzraw <- as.character(peds$genhz) 
        peds$taxonname <- factor(peds$taxonname)
        
        # fix for single-group sets 
        gno <-  c(-6,-10)
        if(length(levels(peds$taxonname)) == 1)
          gno <- -8
        
        groupedProfilePlot(peds, name='genhzraw', 
                           groups = 'taxonname', label='pedon_id',
          print.id = TRUE, id.style = 'side', cex.id=1.2,
          cex.names=1, cex.depth.axis=1.25,y.offset=7,
          axis.line.offset=-3.0, group.name.cex=1,
          group.name.offset = gno, color=input$thematic_field,
          width=0.1, shrink=T, shrink.cutoff=3)
    
        options=list(
          width="100%", height=700
        )
      }
    }
  )
```

### Generalized Hz Probability
```{r, warning=FALSE, echo=FALSE, error=FALSE}
if(loaded)
  renderPlot(
    {
      peds <- getPedonsByPattern(input, pedons, 
                                 input$s.mu,
                                 input$pedon_pattern, 
                                 input$upid_pattern,
                                 input$pedon_list,
                                 input$taxon_kind,
                                 input$phase_pattern)
      
      peds$genhz <- factor(peds$genhz, 
                           levels = guessGenHzLevels(peds)$levels)
      
      if(length(peds)) {
        s <- suppressWarnings(summarize.component(peds))
        return(s$ml.hz.plot)
      }
    }
  )
```

### Depth Class
```{r, warning=FALSE, echo=FALSE, error=FALSE}
if(loaded)
  renderPlot({
    peds <- getPedonsByPattern(input, pedons, 
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
    
    if(length(peds)) {
      depthz <- profileApply(peds, estimateSoilDepth, p="Cr|Cd|R|qm")
      
      return({plot(density(depthz, na.rm=TRUE),
                          main="Density plot of depth to Cr, Cd, R, or *qm horizon",
                          xlab="Depth, centimeters")
              abline(v=c(25,50,100,150))
        })
    }
  })
```

### Depth Table

```{r, warning=FALSE, echo=FALSE, error=FALSE}
if(loaded)
  renderDataTable( {
    peds <- getPedonsByPattern(input, pedons, 
                             input$s.mu,
                             input$pedon_pattern, 
                             input$upid_pattern,
                             input$pedon_list,
                             input$taxon_kind,
                             input$phase_pattern)
  
    if(length(peds)) {
      dc <- aqp::getSoilDepthClass(peds, p="Cr|Cd|R|qm")$depth.class
      dc <- factor(as.character(dc), 
                   levels = c("very.shallow","shallow",
                              "mod.deep","deep","very.deep"),
                   labels = c("very shallow","shallow",
                              "moderately deep","deep","very deep"))
      dct <- as.data.frame(base::table(dc))
      names(dct) <- c("Depth Class", "# of Profiles")
      DT::datatable(dct, options = list(
        bPaginate = FALSE
        
      ))
    }
  }, striped=TRUE, rownames = FALSE)
```

### Geomorphology

#### Hillslope position (2D)
```{r, warning=FALSE, echo=FALSE, error=FALSE}
if(loaded)
  renderTable( {
    
    if(!length(input) | !length(pedons))
      return(data.frame())
    
    peds <- getPedonsByPattern(input, pedons, 
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
    
    if(length(peds)) {
      return(categorical.prop.table(peds$hillslopeprof))
    }
  }, striped=T)

if(loaded)
  renderUI( {
    sourcemu <- input$s.mu 
    peds <- getPedonsByPattern(input, pedons, 
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
    
    if(length(peds) & length(input$modal_pedon) & !is.na(input$modal_pedon)) {
      modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,
                                                     site(peds)$pedon_id,
                                                     sep=":")),] 
      print(paste0("Modal pedon value: ", modal$hillslopeprof))
    }
  })
```

#### Geomorphic position - Hills (3D)
```{r, warning=FALSE, echo=FALSE, error=FALSE}
if(loaded)
  renderTable( {
    peds <- getPedonsByPattern(input, pedons,  
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
    
    if(length(peds) & any(!is.na(peds$geomposhill))) {
      categorical.prop.table(peds$geomposhill) 
    } else {
      print("No Hill 3D Geomorph")
    }
  }, striped=T)

if(loaded) 
  renderUI( {
    peds <- getPedonsByPattern(input, pedons, 
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
    
    if(length(peds) & length(input$modal_pedon) & !is.na(input$modal_pedon)) {
      modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,
                                                     site(peds)$pedon_id,
                                                     sep=":")),] 
      print(paste0("Modal pedon value: ", modal$geomposhill))
    }
  })
```

#### Geomorphic position - Mountains (3D)
```{r, warning=FALSE, echo=FALSE, error=FALSE}
if(loaded)
  renderTable( {
    peds <- getPedonsByPattern(input, pedons, 
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
    if(length(peds) & any(!is.na(peds$geomposmntn))) {
      categorical.prop.table(peds$geomposmntn) 
    } else {
      print("No Mountain 3D Geomorph")
    }
  }, striped=T)

if(loaded)  
  renderUI( {
    peds <- getPedonsByPattern(input, pedons, 
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
    
    if(length(peds) & !is.na(input$modal_pedon)) {
      modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,
                                                     site(peds)$pedon_id,
                                                     sep=":")),] 
      print(paste0("Modal pedon value: ",modal$geomposmntn))
    }
  })
```

#### Geomorphons
```{r, warning=FALSE, echo=FALSE, error=FALSE}
if(loaded)
  renderTable( {
    peds <- getPedonsByPattern(input, pedons, 
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
  
    if(length(peds) & ("gis_geomorphons" %in% siteNames(peds))) {  
      return(categorical.prop.table(peds$gis_geomorphons))
    } else {
      print("Could not find attribute `gis_geomorphons` in `pedon` object.")
    }
  }, striped=T)

if(loaded) 
  renderUI( {
    peds <- getPedonsByPattern(input, pedons, 
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
    
      if(length(peds) & length(input$modal_pedon) & !is.na(input$modal_pedon)) {
        modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,                                                site(peds)$pedon_id,sep=":")),]
        if("gis_geomorphons" %in% siteNames(modal)) {   
          print(paste0("Modal pedon value: ", modal$gis_geomorphons))
        } else {
          print("Could not find attribute `gis_geomorphons` in selected modal pedon")
        }
      }
  })
```

#### Drainage class
```{r, warning=FALSE, echo=FALSE, error=FALSE}
if(loaded)
  renderTable( { 
      peds <- getPedonsByPattern(input, pedons, 
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
      if(length(peds) & any(!is.na(peds$drainagecl))) {
        return(categorical.prop.table(peds$drainagecl))
      }
    }, striped=T)

if(loaded)  
  renderUI( {
    peds <- getPedonsByPattern(input, pedons, 
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
    
    if(length(peds) & length(input$modal_pedon) & !is.na(input$modal_pedon)) {
      modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,
                                                     site(peds)$pedon_id,
                                                     sep=":")),] 
      print(paste0("Modal pedon value: ",modal$drainagecl))
    }
  })
```

#### Surface Shape (DOWN/ACROSS)
```{r, warning=FALSE, echo=FALSE, error=FALSE}
if(loaded)
  renderTable({
    peds <- getPedonsByPattern(input, pedons, 
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
    if(length(peds)) {  
      # make combined curvature classes
      down <- factor(as.character(peds$shapedown),
                     labels =  c("V","L","C"),
                     levels=c("convex","linear","concave"))
      
      acro <- factor(as.character(peds$shapeacross),
                     labels =  c("V","L","C"),
                     levels=c("convex","linear","concave"))
      shape <- factor(paste(as.character(down),
                            as.character(acro),
                            sep="/"))
      
      shape[grepl(shape,pattern="NA")] <- NA
      shape <- factor(shape)
      
      return(categorical.prop.table(shape))
    }
  }, striped=T)

if(loaded)
  renderUI( {
    peds <- getPedonsByPattern(input, pedons, 
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
    
    if(length(peds) & length(input$modal_pedon) & !is.na(input$modal_pedon)) {
      modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,
                                                     site(peds)$pedon_id,
                                                     sep=":")),] 
      print(paste0("Modal pedon value: ",
                   modal$shapedown,
                   modal$shapeacross))
    }
  })
```

### Aspect
```{r, warning=FALSE, echo=FALSE, error=FALSE}
if(loaded)
  renderPlot({
    peds <- getPedonsByPattern(input, pedons, 
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
    
    # cannot make aspect plot with 3 or less profiles
    if(length(peds) & length(peds) > 2)
      aspect.plot(peds$aspect_field, q=p.low.rv.high,
                  plot.title=input$pedon_pattern, pch=21,
                  bg='RoyalBlue', col='black', cex=0.75,
                  arrow.col=c('darkgrey', 'blue', 'darkgrey'))
    else return(-1)
  })

if(loaded)
  renderUI( {
    peds <- getPedonsByPattern(input, pedons, 
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
    
    if(length(peds) & length(input$modal_pedon) & !is.na(input$modal_pedon)) {
      modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,
                                                     site(peds)$pedon_id,
                                                     sep=":")),] 
      print(paste0("Modal pedon value: ", modal$aspect_field))
    }
  })
```

### Ecology

#### Ecological site
```{r, warning=FALSE, echo=FALSE, error=FALSE}
 if(loaded)
  renderTable( { 
    peds <- getPedonsByPattern(input, pedons,
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
    
    df <-  categorical.prop.table(peds$ecositeid)
    df
  }, striped=T)
  
  renderUI( {
    sourcemu <- 
    peds <- getPedonsByPattern(input, pedons, 
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
    
    if(length(peds) & length(input$modal_pedon) & !is.na(input$modal_pedon)) {
      modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,
                                                     site(peds)$pedon_id,
                                                     sep=":")),] 
      print(paste0("Modal pedon value: ", modal$ecositeid))
    }
  })
```

#### Plant list
TODO: Print vegplot data

### Horizon

#### Field-described versus Generalized Horizonation
```{r, warning=FALSE, echo=FALSE, error=FALSE} 
if(loaded)
  renderDataTable({
    peds <- getPedonsByPattern(input, pedons, 
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
    if(length(peds)) {
      m <- as.matrix(table(peds$hzname,peds$genhz))
      colz <- colnames(m)
      rowz <- rownames(m)
      x <- length(colz)
      y <- length(rowz)
      buf <- data.frame(field.hz=rowz)
      for(i in 1:x) {
        buf <- cbind(buf, data.frame(m[,i]))
      }
      colnames(buf) <- c("Field HZ", colz)
      rownames(buf) <- NULL
      DT::datatable(buf, options = list(
        pageLength = 10
      ))
    }
  })
```

### Horizon Relation
```{r, warning=FALSE, echo=FALSE, error=FALSE}
if(loaded)
  renderPlot({
    peds <- getPedonsByPattern(input, pedons,
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern) 
    peds$genhz <- factor(peds$genhz, levels = guessGenHzLevels(peds)$levels)
    if(length(peds)) { 
      # convert contingency table -> adj. matrix
      m <- genhzTableToAdjMat(table(peds$genhz, peds$hzname))
      if(nrow(m))
        plotSoilRelationGraph(m, graph.mode = 'directed', 
                            edge.arrow.size=0.5, 
                            vertex.label.family='sans')
    }
  })
```

### Modal

```{r, warning=FALSE, echo=FALSE, error=FALSE}
#### Modal pedon (field horizonation : generalized horizonation)
if(loaded)
  renderPlot({
    peds <- getPedonsByPattern(input, pedons, 
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
    
    if(length(peds) & length(input$modal_pedon) & !is.na(input$modal_pedon)) {
      
      modal <- peds[which(input$modal_pedon == paste(site(peds)$peiid,
                                                     site(peds)$pedon_id,
                                                     sep=":")),] 
      modal$hzagg <- paste0(modal$hzname,":",modal$genhz)
      
      if(length(modal))
        groupedProfilePlot(modal, name='hzagg',
                         groups="taxonname", label='pedon_id', 
                         print.id = TRUE, id.style = 'side', cex.id=1.2, 
                         cex.names=1, cex.depth.axis=1.25,y.offset=7, 
                         axis.line.offset=-3.0, group.name.cex=1, 
                         group.name.offset = -6,
                         color=input$thematic_field,
                         width=0.1,shrink=T,shrink.cutoff=3)
      options=list(
        width="100%", height=700
      )
    }
  })
  ```
  
  ### Texture
  ```{r, warning=FALSE, echo=FALSE, error=FALSE}
  renderTable({ 
    peds <- getPedonsByPattern(input, pedons, 
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
    
    if(length(peds)) { 
      s <- suppressWarnings(summarize.component(peds))
      s$tt
    }
  })
```

### Texture Triangle
```{r, warning=FALSE, warning=FALSE, results='asis'}
if(loaded)
  renderPlot({
    peds <- getPedonsByPattern(input, pedons, 
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
    if(length(peds)) {
      x <- na.omit(data.frame(genhz=peds$genhz, sand=peds$sand, silt=peds$silt, clay=peds$clay))
      
      if(nrow(x)) {
        x <- x[rowSums(x[,2:4]) == 100,]
        
        aqp::textureTriangleSummary(x[,2:4],
                                    pch=".",
                                    range.col='darkgreen')
        # x$genhz <- factor(x$genhz)
        # # try to set up table structure
        # res <- lapply(split(x, x$genhz), function(hz) {
        #   if(nrow(hz) >= 3) {
        #     aqp::textureTriangleSummary(hz[,2:4], 
        #                                 pch=".", 
        #                                 range.col='darkgreen',
        #                                 main=sprintf("Textures (%s)",
        #                                              unique(hz$genhz)))
        #   }
        # })
        # do.call('rbind', res)
      }
    }
  })
```

### Color
```{r, warning=FALSE, echo=FALSE, error=FALSE}
if(loaded)
  renderPlot({
    peds <- getPedonsByPattern(input, pedons, 
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
    peds$genhz <- factor(peds$genhz, levels = guessGenHzLevels(peds)$levels)
    if(length(peds)) {
      aggregateColorPlot(aggregateColor(peds, groups = 'genhz', 
                                      col = 'soil_color'),
                       label.font = 2, label.cex = 0.95, print.n.hz = TRUE)
    }
  })
```

### Morphology

Numeric attributes summarized by: min, 5th-50th-95th percentiles, max and aggregated by generalized horizon label (NASIS `phorizon` "Component Layer ID"").

```{r, warning=FALSE, echo=FALSE, error=FALSE}
if(loaded)
  
  renderTable({
    peds <- getPedonsByPattern(input, pedons,
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
    peds$genhz <- factor(peds$genhz, levels = guessGenHzLevels(peds)$levels)
    if(length(peds)) {
      s <- suppressWarnings(summarize.component(peds))
      return(s$rt)
    }
  })
```

### Surface Fragments
```{r, warning=FALSE, echo=FALSE, error=FALSE}
if(loaded)
  renderTable({
    peds <- getPedonsByPattern(input, pedons,
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
    if(length(peds)) {
      s <- suppressWarnings(summarize.component(peds))
      return(s$sf)
    }
  })

```

### Diagnostics
```{r, warning=FALSE, echo=FALSE, error=FALSE}
if(loaded)
  renderTable({
    peds <- getPedonsByPattern(input, pedons, 
                               input$s.mu, 
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
    if(length(peds)) {
      s <- summarize.component(peds)
      s$dt
    }
  })
```

### Diagnostics plot
```{r, warning=FALSE, echo=FALSE, error=FALSE}
if(loaded)
  renderPlot({
    peds <- getPedonsByPattern(input, pedons,
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
    
    if(length(peds) & nrow(diagnostic_hz(peds))) {
      #return(diagnostic_hz(peds))
      v.possible <- c("anthropic epipedon","abrupt textural change",
              "andic soil properties","cambic horizon","calcic horizon","durinodes",
              "petrocalcic horizon",
             "fragipan","gypsic horizon","gilgai","glossic horizon","aquic conditions",
             "histic epipedon","sombric horizon","petrogypsic horizon","placic horizon",
             "kandic horizon","lithic contact","secondary carbonates","mollic epipedon",
             "melanic epipedon","natric horizon","ochric epipedon","plaggen epipedon",
             "petroferric contact","permafrost","plinthite","albic horizon",
             "agric horizon","spodic horizon","slickensides","argillic horizon",
             "umbric epipedon","sulfuric horizon","paralithic contact","oxic horizon",
             "salic horizon","duripan","sulfidic materials",
             "interfingering of albic materials","fibric soil materials",
             "hemic soil materials","sapric soil materials","humilluvic material",
             "limnic materials","coprogenous earth","diatomaceous earth","marl",
             "albic materials","mottles with chroma 2 or less","lamellae",
             "ortstein","fragic soil properties","densic contact","densic materials",
             "paralithic materials","anhydrous conditions","cryoturbation",
             "gelic materials","glacic layer","folistic epipedon",
             "redox concentrations","redox depletions with chroma 2 or less",
             "reduced matrix","endosaturation","episaturation","anthric saturation",
             "lithologic discontinuity","strongly contrasting particle size class",
             "gypsum accumulations","salt accumulations","volcanic glass",
             "n value > 0.7","spodic materials","artifacts","fibers",
             "free carbonates","resistant minerals","weatherable minerals",
             "anhydritic horizon","human-altered material","human-transported material",
             "manufactured layer","manufactured layer contact")
      v.use <- gsub(v.possible, pattern = "\\s{1}", replacement=".")
      v <- v.use[v.use %in% siteNames(peds)]
      
      h <- site(peds)
      lapply(as.list(v), function(vv) {
        h[is.na(h[,vv]), vv] <- FALSE 
      })
      diagnosticPropertyPlot(peds, v = v, k = 2)
    } else {
      print("No pedons matching criteria or diagnostic features populated.")
    }
  })
```

### Mapunit Summary
```{r, warning=FALSE, echo=FALSE, error=FALSE}
if(loaded)
  renderPlot({
    peds <- getPedonsByPattern(input, pedons,
                               input$s.mu,
                               input$pedon_pattern, 
                               input$upid_pattern,
                               input$pedon_list,
                               input$taxon_kind,
                               input$phase_pattern)
    
    if(length(peds)) {
      s <- suppressWarnings(summarize.component(peds))
      print(s$pmg)
    }
  })
```