---
title: Summary of NASIS NCSS Pedon Lab Data Table
author: "`r Sys.getenv('USERNAME')`"
date: "`r Sys.Date()`"
output:
  html_document:
    keep_md: no
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
params:
   series: !r as.character("Miami")
   genhz_rules: !r as.character(file.path(system.file(package = "soilReports"), "reports/northcentral/lab_summary_by_taxonname/genhz_rules/Generic_rules.R"))
editor_options: 
  chunk_output_type: console
---
```{r setup, echo=FALSE, warning=FALSE}
p <- c(0, 0.25, 0.5, 0.75, 1)

# knit options
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, background="#F7F7F7", fig.retina=1, dev="png", tidy=FALSE, verbose=FALSE, results = 'asis')

options(stringsAsFactors = FALSE)
```


```{r load packages, include=FALSE}
source(params$genhz_rules)

# soil libraries
library(aqp)
library(soilDB)
library(soilReports)

# data manipulation libraries
library(knitr)
library(plyr)
library(reshape2)
library(circular)

# graphic libraries
library(lattice)
library(latticeExtra)
library(RColorBrewer)

# mapping libraries
library(maps)
library(maptools)
library(mapview)

# custom ggplot2like latticeExtra theme
opar <- trellis.par.get()
trellis.par.set(theme = ggplot2like())
tpg <- trellis.par.get()
tpg$axis.line <-  list(alpha = 1, col = "#000000", lty = rep(1, 7), lwd = rep(0.2, 7))
tpg$strip.border <- list(alpha = 1, col = "#000000", lty = rep(1, 7), lwd = rep(0.2, 7))
trellis.par.set(tpg)
```

```{r fetch and format}
# load NASIS data
l <- fetchNASISLabData()
f <- fetchNASIS()

lh <- horizons(l)
lp <- site(l)
h <- horizons(f)
s <- site(f)

# merge tables
lh_names <- names(lh)
h_vars <- c("labsampnum", "hzdept", "hzdepb", "fragvoltot", "total_frags_pct")
lp_vars <- c("peiid", "labpeiid")
s_vars <- c("peiid", "earthcovkind1")

lh <- merge(lh, h[h_vars], by = c("labsampnum", "hzdept", "hzdepb"), all.x = TRUE, sort = FALSE)
lh <- merge(lh, lp[lp_vars], by = "labpeiid", all.x = TRUE, sort = FALSE)
lh <- merge(lh, s[s_vars], by = "peiid", all.x = TRUE, sort = FALSE)

lh <- lh[order(lh$labpeiid), c(lh_names, h_vars[-1:-3], s_vars[-1])]

# transform variables
lh <- within(lh, {
  fragvoltot = ifelse(is.na(fragvoltot), total_frags_pct, fragvoltot)
  hzname = ifelse(is.na(hzname), hznameoriginal, hzname)
  hzname[is.na(lh$hzname)] = "missing"
  thickness = hzdepb - hzdept
  hzdepm = (hzdept + hzdepb) / 2
  genhz = generalize.hz(hzname, ghr$n, ghr$p, hzdepm = hzdepm)
  earthcovkind1[!grepl("^A", genhz)] = ""
  genhz_nlcd = ifelse(grepl("^A", genhz), 
                      paste(genhz, abbreviate(earthcovkind1, 10), sep = " - "),
                      as.character(genhz)
                      )
  genhz_nlcd = factor(genhz_nlcd,
                      levels = names(sort(tapply(hzdepm, genhz_nlcd, median)))
                      )
  awc = round((wthirdbarclod - wfifteenbar) * 
                dbthirdbar * 
                (1 - total_frags_pct / 100) / 100, 2)
  })

horizons(l) <- lh

lp <- within(lp, {
  cec7clayratiowtavg = cec7clayratiowtavg * 100
  thickness = pscbotdepth - psctopdepth
  })

h <- within(h, {
  hzname[is.na(h$hzname)] = "missing"
  hzdepm = (hzdept + hzdepb) / 2
  genhz = generalize.hz(hzname, ghr$n, ghr$p, hzdepm = hzdepm)
  })
horizons(f) <- h


# indexing info
lh_i <- c('peiid', 'labpeiid', 'labsampnum', 'layerseqnum', 'hzdept', 'hzdepb', 'hzdepm', 'layertype', 'hzname', 'hznameoriginal', 'genhz', 'moist_prepstate', 'coverkind_1', 'recwlupdated', 'ncsslayerlabdataiid')

# static soil properties
lh_s <- c('texcl', 'stratextsflag', 'sandvc', 'sandco', 'sandmed', 'sandfine', 'sandvf', 'sandtot', 'siltco', 'siltfine', 'silttot', 'claycarb', 'clayfine', 'claytot', 'ompctest', 'fiberrubbedpct', 'fiberunrubbedpct', 'fragwt25', 'fragwt520', 'fragwt2075', 'fragwt275', 'wtpct0175', 'wtpctgt2ws')

# dynamic soil properties
lh_d <- c('organicmatpct', 'carbonorganicpct', 'carbontotalpct', 'ph1to1h2o', 'ph01mcacl2', 'phnaf', 'phoxidized', 'resistivity', 'ec', 'esp', 'sar', 'cecsumcations', 'cec7', 'ecec', 'sumbases', 'basesatsumcations', 'basesatnh4oac', 'caco3equiv', 'caco3lt20', 'gypsumequiv', 'feoxalate', 'feextractable', 'fetotal', 'sioxalate', 'extracid', 'extral', 'aloxalate', 'altotal', 'pmehlich3', 'ph2osoluble', 'poxalate', 'polsen', 'ptotal', 'nzpretention', 'dbthirdbar', 'dbovendry', 'aggstabpct', 'wtenthbarclod', 'wtenthbarsieve', 'wthirdbarclod', 'wthirdbarsieve', 'wfifteenbar', 'wretentiondiffws', 'wfifteenbartoclay', 'awc', 'adod', 'lep', 'cole', 'liquidlimit', 'pi', 'cec7clay')

# fig.height
ssp_nh <- ceiling(length(lh_s[lh_s %in% names(lh)]) / 4) * 3
dsp_nh <- ceiling(length(lh_d[lh_d %in% names(lh)]) / 4) * 3
dsp_nh <- ifelse(dsp_nh < 1, 1, dsp_nh)
```


# Brief summary of NCSS lab pedon data

## Interactive Map

```{r map, echo=FALSE, results='asis'}

if (dim(s)[1] != 0) {
  pedon_locations <- s[complete.cases(s[c("x_std", "y_std")]), ]
  coordinates(pedon_locations) <- ~ x_std + y_std
  proj4string(pedon_locations) <- CRS("+init=epsg:4326")
  
  if(params$series != "Generic") {
    series_extent <- seriesExtent(params$series)
    
    mapView(pedon_locations) + series_extent
    } else mapView(pedon_locations)
  } else("no coordinates")
```

## Site Data

```{r site data}
kable(s[, c("upedonid", "taxonname", "taxsubgrp", "taxpartsize", "pedontype", "describer")], cation = "Site Data")
```


## Comparison of genhz and hzname pattern matching

```{r genhz matching, fig.width=8}
hz_t <- addmargins(table(lh$genhz, lh$hzname))

idx <- pIndex(hz_t, 15)

for (i in unique(idx)){
  print(kable(hz_t[, c(idx == i)], align = "c", digits = 0, caption = "Horizon designations vs generic horizon designations (counts)"))
  }
```


## Profile Plots

```{r profile plots, eval=FALSE}
cols <- brewer.pal(n = length(levels(l$genhz)), name = "Set1") 
hz.names <- levels(l$genhz)
l$genhz.soil_color <- cols[match(l$genhz, hz.names)] # assign a color to each generalized horizon label

idx <- pIndex(l, 15)

# plot no more than 15 soil profiles on each row
for (i in unique(idx)){
  plot(l[which(idx == i)], name = 'hzname', color = "genhz.soil_color", label = 'upedonid')
  title(" Soil Profile Plots")
  legend('bottomleft', legend = hz.names, pt.bg = cols, pch = 22, horiz = TRUE, pt.cex = 2, text.width = 1)
  }
```


# Range in characteristics (RIC) for NCSS pedon lab data

## Generalized horizon depths

```{r}
lh.lo <- melt(lh, id.vars = "genhz", measure.vars = c("hzdept", "hzdepb", "thickness"))
lh.5n <- ddply(lh.lo, .(variable, genhz), summarize,
               range = prettySummary(value)
               )
lh.c <- dcast(lh.5n, genhz ~ variable, value.var = "range")

kable(lh.c, digits = 0, align = "c", caption = "Depths and thickness of generic horizons (min, 25th, median, 75th, max)(n)")

lh.lo$genhz <- factor(lh.lo$genhz, levels = rev(levels(lh.lo$genhz)))

bwplot(genhz ~ value | variable, data = lh.lo, 
       main = "Box plots of numeric variables by generic horizon", 
       scales =list(x="free"), axis = axis.grid,
       as.table = TRUE
       )
```


## Aggregate soil profile attributes

Five number summary (min, 25th, median, 75th, max)(percentiles)

```{r aggregate measures, fig.width=8}
# Summarize site data
vars <- c("noncarbclaywtavg", "claytotwtavg", "le0to100", "wf0175wtavgpsc", "volfractgt2wtavg", "cec7clayratiowtavg")
lp.sub <- lp[vars]
lp.lo <- melt(lp.sub, measure.vars = vars)
lp.5n <- ddply(lp.lo, .(variable), summarize, 
               range = prettySummary(value)
               )

kable(lp.5n, caption = "Aggregate soil profile attributes", align = "c")

if (nrow(na.exclude(lp.lo)) > 6) {
  bwplot(variable ~ value, data = lp.lo, 
         main = "Aggregate Soil Profile Attributes", 
         xlab = "Percent (%)", ylab = "Soil Properties",
         axis = axis.grid,
         as.table = TRUE
         )
  }         

lp.lo <- melt(lp, id.vars="labpeiid", measure.vars = c("psctopdepth", "pscbotdepth", "thickness"))
lp.5n <- ddply(lp.lo, .(variable), summarize, 
               range = prettySummary(value)
               )

kable(lp.5n, digits = 0, align = "c", caption = "Depths and thickness of particle size control section (min, 25th, median, 75th, max)(n)")

if (length(na.exclude(lp$thickness)) > 1){
  bwplot(value ~ variable, data=lp.lo, 
         main = "Particle Size Control Section Depths and Thickness",
         ylab="cm",
         ylim = c(max(lp.lo$value, na.rm = T) + 5, 
                  min(lp.lo$value, na.rm = T) - 5
                  ), 
         scales = list(x = "free"), axis = axis.grid
         )
  }
```


## Range in characteristics (RIC) for static soil properties 

### Tables

Five number summary (min, 25th, median, 75th, max)(percentiles) and total number of observations (n)

```{r ssp, fig.height=ssp_nh, fig.width=9}

# Tables

lh.num <- lh[c("genhz", lh_s[-c (1:2)])]
lh.num <- Filter(f = function(x) !all(is.na(x)), x = lh.num)
lh.lo <- melt(lh.num, id.vars = "genhz")
lh.5n <- ddply(lh.lo, .(variable, genhz), summarize,
               range = prettySummary(value)
               )
lh.c <- dcast(lh.5n, genhz ~ variable, value.var = "range")

idx <- pIndex(lh.c, 4)

for (i in unique(idx)){
  print(kable(lh.c[, c(TRUE, idx == i)], align = "c"))
  }

# Boxplots

lh.lo$genhz <- factor(lh.lo$genhz, levels = rev(levels(lh.lo$genhz)))
n <- ceiling(length(levels(lh.lo$variable))/4)
```

</p>


### Boxplots

```{r ssp boxplots, fig.height=ssp_nh, fig.width=9}

bwplot(genhz ~ value | variable, data = lh.lo,
       main = "Box plots of static soil properties by generic horizon", 
       scales=list(x = "free"), axis = axis.grid, 
       as.table = TRUE, layout = c(4, n)
       )
```


## Range in characteristics (RIC) for dynamic soil properties 

### Tables

```{r dsp}

# Tables

lh.num <- lh[c("genhz_nlcd", lh_d)]
lh.num <- lh.num[apply(lh.num, 1, function(x) !all(is.na(x))), ]
lh.lo <- reshape2::melt(lh.num, id.vars = "genhz_nlcd")
lh.5n <- ddply(lh.lo, .(variable, genhz_nlcd), summarize, 
               range = prettySummary(value)
               )
lh.c <- reshape2::dcast(lh.5n, genhz_nlcd ~ variable, value.var = "range")

idx <- pIndex(lh.c, 4)

for (i in unique(idx)){
  print(knitr::kable(lh.c[, c(TRUE, idx == i)], align = "c"))
  }
```

</p>

### Boxplots

```{r dsp boxplots, fig.height=dsp_nh, fig.width=9}
# Boxplots

lh.lo$genhz_nlcd <- with(lh.lo, factor(genhz_nlcd, levels = rev(levels(genhz_nlcd))))

n <- ceiling(length(levels(lh.lo$variable))/4)

bwplot(genhz_nlcd ~ value | variable, data = lh.lo,
       main = "Box plots of dynamic soil properties by generic horizon", 
       scales=list(x = "free"), axis = axis.grid, 
       as.table = TRUE, layout = c(4, n)
       )
```


## Texture

```{r texture}
kable(addmargins(xtabs(~ genhz + texcl, data = lh, drop.unused.levels = TRUE)), digits = 0, caption = "Texture by generic horizon")
```
