---
title: Summary of NASIS Pedon Horizon Table
date: "`r Sys.Date()`"
output:
  html_document:
    keep_md: yes
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
---

```{r report-metadata, echo=FALSE, results='hide', warning=FALSE, message=FALSE, include=FALSE}
## version number
.report.version <- 1.0

## short description
.report.description <- 'summarize field pedons from NASIS pedon table'
```

```{r setup, echo=FALSE, warning=FALSE}
# report parameters
source("config.R")

# custom functions 
source("utils.R")

# generic horizon rules
source(paste0("./genhz_rules/", series, "_rules.R"))

# knit options
knitr::opts_chunk$set(echo=FALSE, results='asis', warning=FALSE, message=FALSE, background="#F7F7F7", fig.retina=1, dev="png", tidy=FALSE, verbose=FALSE)

options(stringsAsFactors = FALSE)
```

```{r current environment}
# save current workspace image and reload, for some reason when a report is knit using render() doesn't clear the temporary objects from the global environment
save.image(file = "C:/Temp/temp.RData")
```

```{r load packages}
# soil libraries
library(aqp)
library(soilDB)

# data manipulation libraries
library(knitr)
library(plyr)
library(reshape2)
library(circular)

# graphic libraries
library(lattice)
library(latticeExtra)
library(RColorBrewer)

# mapping libraries
library(maps)
library(maptools)
library(mapview)

# custom ggplot2like latticeExtra theme
opar <- trellis.par.get()
trellis.par.set(theme = ggplot2like())
tpg <- trellis.par.get()
tpg$axis.line <-  list(alpha = 1, col = "#000000", lty = rep(1, 7), lwd = rep(0.2, 7))
tpg$strip.border <- list(alpha = 1, col = "#000000", lty = rep(1, 7), lwd = rep(0.2, 7))
trellis.par.set(tpg)
```


```{r fetch and format, load-data, echo=FALSE, warning=FALSE}
# load NASIS data
pedons <- fetchNASIS_pedons(lab = TRUE)

h <- horizons(pedons)
s <- site(pedons)
d <- diagnostic_hz(pedons)

# modify the horizon table
h <- within(h, {
  hzname = ifelse(is.na(hzname), "missing", hzname)
  hzdepm = (hzdept + hzdepb) / 2
  genhz  = generalize.hz(hzname, ghr$n, ghr$p, hzdepm = hzdepm)
  gravel = gravel - fine_gravel
  })
var <- c("fine_gravel", "gravel", "cobbles", "stones", "boulders", "paragravel", "paracobbles", "channers", "flagstones")
h[var] = lapply(h[var], as.integer)
  
horizons(pedons) <- h

h <- data.frame(lapply(h, na_replace), stringsAsFactors = FALSE)
s <- data.frame(lapply(s, na_replace), stringsAsFactors = FALSE)

# rename rock fragments
idx <- names(h) == "total_frags_pct"
names(h)[idx] <- c("fragvoltotc")

# vector of names to exclude from numeric summary
vars <- c('peiid', 'phiid', 'hzname', 'genhz', 'hzdept', 'hzdepb', 'labsampnum', 'd_value', 'd_chroma', 'm_value', 'm_chroma', 'parafine_gravel', 'paragravel', 'paracobbles', 'parastones', 'paraboulders', 'parachanners', 'paraflagstones', 'unspecified', 'd_r', 'd_b', 'd_g', 'd_sigma', 'm_r', 'm_b', 'm_g', 'm_sigma')

# create vector of names for numeric soil variables excluding vars vector
num_vars <- names(h)[! names(h) %in%  vars &                     # remove vars
                       sapply(h, is.numeric) &                   # select numeric columns
                       sapply(h, function(x) !all(is.na(x)))  &  # select non-empty columns 
                       sapply(h, function(x) !all(x == 0))       # remove columns were all observations == 0
                     ]

# fig.height
nh <- length(num_vars) / 4 * 2.5

# modify diagnostic table
# create a empty diagnostic table if data is NULL
if (nrow(d) == 0) {
  d <- data.frame(peiid = s$peiid,
                  diag_kind = as.character("missing"),
                  featdept = as.integer(NA), 
                  featdepb = as.integer(NA)
                  )
  }
d <- transform(d, thickness = featdepb - featdept)

# modify site table
# rename psc depths
idx <- names(s) == "psctopdepth" | names(s) == "pscbotdepth"
names(s)[idx] <- c("featdept", "featdepb")

s <- transform(s, thickness = featdepb - featdept)

srf <- s[grepl("surface_", names(s))]
names(srf) <- gsub("surface_", "", names(srf))
srf <- within(srf, {
  total_srf = gravel + cobbles + stones + boulders + flagstones + channers
  gravel = gravel - fgravel
  })
```

# Brief summary of pedon data

## Interactive Map

```{r map, echo=FALSE, results='asis'}

if (dim(s)[1] != 0) {
  pedon_locations <- s[complete.cases(s[c("x_std", "y_std")]), ]
  coordinates(pedon_locations) <- ~ x_std + y_std
  proj4string(pedon_locations) <- CRS("+init=epsg:4326")
  
  if(series != "Generic") {
    series_extent <- seriesExtent(series)
    
    mapView(pedon_locations) + series_extent
    } else mapView(pedon_locations)
  } else("no coordinates")
```

## Site Data

```{r format site data, results='asis', echo=FALSE}
# Site information
kable(subset(s, select = c("pedon_id", "taxonname", "tax_subgroup", "part_size_class", "pedon_type", "describer")), caption = "Summary of data in the selected set")
```


## Comparison of genhz and hzname pattern matching

```{r genhz matching, fig.width=8}
hz_t <- addmargins(table(h$genhz, h$hzname))

idx <- pindex(hz_t, 15)

# plot 15 horizon designations per row
for (i in unique(idx)){
  print(kable(hz_t[, c(idx == i)], align = "c", digits = 0, caption = "Horizon designations vs generic horizon designations (counts)"))
  }
```


## Profile Plots

```{r, echo=FALSE}
cols <- brewer.pal(n = length(levels(pedons$genhz)), name = "Set1") 
hz.names <- levels(pedons$genhz)
# assign a color to each generalized horizon label
pedons$genhz.soil_color <- cols[match(pedons$genhz, hz.names)]

idx <- pindex(pedons, 15)

# plot 15 profiles at a time
for (i in unique(idx)){
  plot(pedons[idx == i], name = 'hzname', color = 'genhz.soil_color', label = 'pedon_id')
  title("Soil profile plots")
  legend('bottomleft', legend = hz.names, pt.bg = cols, pch = 22, horiz = TRUE, pt.cex = 2, text.width = 1)
  }
```


# Range in characteristics (RIC) for NCSS pedon lab data

## Surface rock fragments

```{r surface rock fragments, echo=FALSE, results='asis'}
vars <- c("total_srf", "fgravel", "gravel", "cobbles", "stones", "boulders", "channers", "flagstones")
srf.lo <- melt(srf, measure.vars = vars)
srf.5n <- ddply(srf.lo, .(variable), .fun = sum5n)

kable(srf.5n, align = "c", caption =  "Surface rock fragments (min, 25th, median, 75th, max)(n)")
 
if (sum(srf$total_srf, na.rm = T) != 0) {
  bwplot(variable ~ value, data = srf.lo, 
         main = "Boxplots of surface rock fragments",
         ylab = "percent",
         axis = axis.grid
         )
  }         
```

## Diagnostic horizons and soil characteristics

```{r diagnostics, echo=FALSE, results='asis', warning=FALSE}
diag.lo <- melt(d, id.vars = "diag_kind", measure.vars = c("featdept", "featdepb", "thickness"))
pscs.lo <- melt(s, id.vars = "peiid", measure.vars = c("featdept", "featdepb", "thickness"))
pscs.lo <- data.frame(diag_kind = "particle size control section", 
                      variable = pscs.lo$variable, 
                      value = pscs.lo$value
                      )
# combine diagnostic and particle size control section long tables
diag.lo <- rbind(diag.lo, pscs.lo)
diag.5n <- ddply(diag.lo, .(variable, diag_kind), .fun = sum5n)
diag.wi <- dcast(diag.5n, diag_kind ~ variable, value.var = 'range')

kable(diag.wi, align = "c", caption = "Depths and thickness of diagnostic horizons and features(min, 25th, median, 75th, max)(n)")


if (!all(is.na(diag.lo$value))) {
  bwplot(diag_kind ~ value | variable, data = diag.lo, 
         main = "Boxplots of diagnostic horizon and feature depths", 
         scales =list(x="free"), axis = axis.grid, 
         as.table = TRUE
         )
  }
```


## Generalized horizon depths

```{r hzname vs genhz, echo=FALSE, results='asis'}
genhz.lo <- melt(h, id.vars="genhz", measure.vars = c('hzdept', 'hzdepb'))
genhz.thk <- ddply(h, .(phiid, genhz), summarize, thickness=sum(hzdepb-hzdept))
genhz.lo2 <- melt(genhz.thk, id.vars = "genhz", measure.vars = 'thickness')
genhz.lo <- rbind(genhz.lo, genhz.lo2)
genhz.5n <- ddply(genhz.lo, .(variable, genhz), .fun = sum5n)

kable(dcast(genhz.5n, genhz ~ variable, value.var = 'range'), align = "c", caption = "Depths and thickness of generic horizons (min, 25th, median, 75th, max)(n)")

genhz.lo$genhz <- factor(genhz.lo$genhz, levels = rev(levels(genhz.lo$genhz)))

bwplot(genhz ~ value | variable, data = genhz.lo, 
       main = "Boxplots of horizon generic horizon depths and thickness",
       scales =list(x="free"), axis = axis.grid,
       as.table = TRUE
       )
```


## Range in characteristics (RIC) for static soil properties 

### Tables

```{r ric ssp tables}
h.lo <- melt(h, id.vars="genhz", measure.vars = num_vars)
h.5n <- ddply(h.lo, .(variable, genhz), .fun = sum5n)
h.wi <- dcast(h.5n, genhz ~ variable, value.var = 'range')

idx <- pindex(h.wi, 5)

for (i in unique(idx)) {
  print(kable(h.wi[, c(T, idx == i)], align = "c", caption = "Numeric variables by generic horizon (min, 25th, median, 75th, max)(n)"))
  # inserting an empty line so the last table doesn't come out corrupted
  cat("\n")
  }

h.lo$genhz <- factor(h.lo$genhz, levels = rev(levels(h.lo$genhz)))
n <- ceiling(length(levels(h.lo$variable))/4)
```

### Boxplots

```{r ric ssp boxplots, fig.height=nh, fig.width=8}
bwplot(genhz ~ value | variable, data = h.lo,
       main = "Box plots of numeric variables by generic horizon",
       scales=list(x="free"), axis = axis.grid,
       as.table = TRUE, layout = c(4, n)
       )
```


### Texture

```{r texture}
kable(addmargins(xtabs(~ genhz + texcl, data = h, drop.unused.levels = TRUE)), digits = 0, caption = "Texture by generic horizon (counts)")

hz_tex <- addmargins(xtabs(~ genhz + texture, data = h))
idx <- pindex(hz_tex, 15)

for (i in unique(idx)){
  print(kable(hz_tex[, c(idx == i)], align = "c", digits = 0, caption = "Tex Mod & Class by generic horizon (counts)"))
  }
```

### Color

```{r color}
kable(addmargins(xtabs(~ h$genhz + h$d_hue, data = h, drop.unused.levels = TRUE)), digits = 0, caption = "Dry hue by generic horizon (counts)")

kable(addmargins(xtabs(~ genhz + m_hue, data = h, drop.unused.levels = TRUE)), digits = 0, caption = "Moist hue by generic horizon (counts)")
```

### Effervescence

```{r effervesence}
kable(addmargins(xtabs(~ genhz + effclass, data = h, drop.unused.levels = TRUE)), digits = 0, caption = "Effervescence by generic horizon (counts)")
```


## Range in characteristics (RIC) for the geographic setting

### Elevation, Slope and Aspect

```{r, echo=FALSE, results='asis', warning=FALSE}
vars <- c("elev_field", "slope_field")
morf <- subset(s, select = vars)
morf.lo <- melt(morf, measure.vars = vars)
morf.5n <- ddply(morf.lo, .(variable), .fun = sum5n)

if (!all(is.na(s$aspect_field))) {
  aspect <- subset(s, select = c("aspect_field"))
  aspect.lo <- melt(aspect, measure.vars = "aspect_field")
  aspect.lo$value <- circular(aspect.lo$value, template="geographic", units="degrees", modulo="2pi")
  aspect.5n <- ddply(aspect.lo, .(variable), .fun = sum5n)
  kable(rbind(morf.5n, aspect.5n), caption = "Elevation, slope gradient and aspect (min, 25th, median, 75th, max)(n)", align = "c")
  } else(kable(morf.5n, caption="Elevation and slope gradient (min, 25th, median, 75th, max)(n)", align = "c"))         

bwplot(~ value | variable, data = morf.lo, 
       main = "Boxplots of elevation and slope gradient",
       scales=list(x="free"), axis = axis.grid,
       as.table = TRUE
       )
```

### Parent Material vs. Landform

```{r parent material}
if (!all(is.na(s[c("pmorigin", "pmkind", "landform.string")]))) {
  pm_comb <- factor(paste0(s$pmorigin, " ", s$pmkind))
  pm_land <- factor(s$landform.string)
  pm.lf <- addmargins(table(pm_comb, pm_land))
  kable(pm.lf, caption="Parent material vs landform (counts)")
  } else "The parent material and landform fields are empty."
```

### Slope Shape

```{r shape}
if (!all(is.na(s[c("shapedown", "shapeacross")]))) {
  kable(addmargins(xtabs(~ shapedown + shapeacross, data = s)), caption = "Down slope (y-axis) vs across slope (x-axis) (counts)")
  } else "The slope shape fields are empty."
```

### Hillslope Position vs. Drainage Class

```{r hillslope position}
if (any(complete.cases(s[c("hillslope_pos", "drainagecl")]))) {
  kable(addmargins(xtabs(~ drainagecl + hillslope_pos, data = s)), digits = 0, caption = "Drainage class vs hillslope position (counts)")
  } else "The hillslope position and drainage class fields are empty."
```
